---
title: "Exhaustion"
author: "LinyiGuo"
date: "2019/10/14"
output: 
  html_document:
    toc: true
  pdf_document:
    toc: true
---

**Preface**

&emsp;&emsp;This file is to determine the expected value of parameters. We are going to compare the results from *TRAMO-SEATS/X11* with those from *SSM* under different parameters. As for how to choose the parameters, we use exhaustion method here.


```{r}
rm(list=ls())
set.seed(9483)
```
```{r}
library(KFAS)
library(seasonal)
library(ggplot2)
```


```{r}
# note: data include the result after seasonal adjustment
total <- read.csv('C:\\Users\\GuoLY\\Desktop\\StatCan\\retailtrade data\\total.csv')
automobile <- read.csv('C:\\Users\\GuoLY\\Desktop\\StatCan\\retailtrade data\\automobile.csv')
```

```{r}
total <- ts(total$unadjusted, start = c(2004,01), frequency = 12)
automobile <- ts(automobile$unadjusted, start = c(2004,01), frequency = 12)
```

```{r}
total_x11 <- seas(total, x11='')
automobile_x11 <- seas(automobile, x11='')

summary(total_x11)
summary(automobile_x11)
```

```{r}
total_x11_seasadj <- series(total_x11, 'd11')
total_x11_trend <- series(total_x11, 'd12')
total_x11_seasonal <- series(total_x11, 'd10')
automobile_x11_seasadj <- series(automobile_x11,'d11')
automobile_x11_trend <- series(automobile_x11, 'd12')
automobile_x11_seasonal <- series(automobile_x11, 'd10')
```

```{r}
par(mfrow=c(2,2))
plot(total)
plot(total_x11_seasadj)
plot(total_x11_trend)
plot(total_x11_seasonal)

plot(automobile)
plot(automobile_x11_seasadj)
plot(automobile_x11_trend)
plot(automobile_x11_seasonal)
par(mfrow=c(1,1))
```


```{r ssm}
total_ssmm <- SSModel(log(total) ~ SSMtrend(1, Q=list(1)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=1), 
                     H=1)

total_ssm <- KFS(total_ssmm)
```

```{r}
total_ssm_trend <- coef(total_ssm, states = 'trend')
total_ssm_seasonal <- ts(-rowSums(coef(total_ssm, states = 'seasonal')),
                       start = c(2004,1), frequency = 12)
```

```{r season}
plot(exp(total_ssm_seasonal)[-length(total)], ylab='',ylim=c(0.8,1.15), type='l')
par(new=TRUE)
plot(total_x11_seasonal[-1], ylab='', ylim=c(0.8,1.15), col=2, type='l')

```
```{r trend}
plot(exp(total_ssm_trend), ylab='', ylim=c(2.5*10^7, 5.5*10^7))
par(new=TRUE)
plot(total_x11_trend, ylab='', ylim=c(2.5*10^7, 5.5*10^7), col=2)
```

```{r Seasadj}
plot(total[-1]/exp(total_ssm_seasonal)[-length(total)], type = 'l',
     ylim=c(2.5*10^7,5.5*10^7), ylab='')
par(new=TRUE)
plot(total_x11_seasadj[-1], col=2, type='l',
     ylim=c(2.5*10^7,5.5*10^7), ylab='')
```

what if I use 'total' after prepocessing in ssm? 

```{r}
total1 <- series(total_x11, 'b1')
total1_ssmm <- SSModel(log(total1) ~ SSMtrend(1, Q=list(1)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=1), 
                     H=1)

total1_ssm <- KFS(total1_ssmm)
```
```{r}
total1_ssm_trend <- coef(total_ssm, states = 'trend')
total1_ssm_seasonal <- ts(-rowSums(coef(total1_ssm, states = 'seasonal')),
                       start = c(2004,1), frequency = 12)


plot(exp(total1_ssm_trend), ylab='', ylim=c(2.5*10^7, 5.5*10^7))
par(new=TRUE)
plot(total_x11_trend, ylab='', ylim=c(2.5*10^7, 5.5*10^7), col=2)

plot(total1[-1]/exp(total1_ssm_seasonal)[-length(total)], type = 'l',
     ylim=c(2.5*10^7,5.5*10^7), ylab='')
par(new=TRUE)
plot(total_x11_seasadj[-1], col=2, type='l',
     ylim=c(2.5*10^7,5.5*10^7), ylab='')
```


**Comment: ** I picked these variances randomly, but the results are not ideal. I'm going to use exhaustion to choose a set of variances according to the difference of seasonal adjusted series.

```{r}
# suppose our model is multiplicative
exhaustion_product <- function(data){
  Dif <- c()
  index <- c()
  x11 <- seas(data, x11='')
  Seasadj_x11 <- series(x11, 'd11')
  for (i in 1:6) {
    for (j in 1:6) {
      for (k in 1:6) {
        ssm <- SSModel(log(data) ~ SSMtrend(1, Q=list(10^(-4+j))) + 
                  SSMseasonal(12, sea.type = 'dummy', Q = 10^(-4+k)),
                H = 10^(-4+i))
        mod <- KFS(ssm)
        
        Season_ssm <- -rowSums(coef(mod, states = 'seasonal'))
        Seasadj_ssm_log <- log(data) - Season_ssm
        Seasadj_ssm <- exp(Seasadj_ssm_log)
        
        dif <- sum((Seasadj_ssm - Seasadj_x11)^2)
        Dif <- c(Dif, dif)
        index <- rbind(index, c(10^(-4+i),10^(-4+j),10^(-4+k)))
      }
      
    }
    
  }
  fit <- index[which.min(Dif),]
  return(fit)
}
```

```{r}
exhaustion_product(total)
```

```{r}
ssm_total <- SSModel(log(total) ~ SSMtrend(1, Q=list(100)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=1), 
                     H=0.001)

mod_total_ssm <- KFS(ssm_total)

plot(exp(-rowSums(coef(mod_total_ssm, states='seasonal'))), type='l')

plot(exp(coef(mod_total_ssm, states = 'trend')))

plot(total/exp(-rowSums(coef(mod_total_ssm, states='seasonal'))), type='l')

```

# See Exhaustion2

