---
title: "Exhaustion3"
author: "LinyiGuo"
date: "2019/10/23"
output: 
  html_document:
    toc: true
  pdf_document:
    toc: true
---


```{r include=FALSE}
rm(list=ls())
set.seed(9483)
```
```{r include=FALSE}
library(seasonal)
library(KFAS)
```

# Part 1

Based on Aaron's suggestion, I first change the standard(error) to the sum of (normalized) different components' squared differences, that is:
$${\sum(X11_{trend}-SSM_{trend})^2} + {\sum(X11_{seasonal}-SSM_{seasonal})^2} + {\sum(X11_{seasadj}-SSM_{seasadj})^2}$$


```{r}
# define our standard to choose parameters
# note this is for additive models
Dif <- function(x11, ssm, data){
  
  x11_trend <- series(x11, 'd12')
  x11_seasonal <- series(x11, 'd10')
  x11_seasadj <- series(x11, 'd11')
  
  ssm_trend <- coef(ssm, states = 'trend')
  ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  ssm_seasadj <- data[-1] - ssm_seasonal[-length(data)] # length is shorter
 
  D <-  sum((x11_trend-ssm_trend)^2) + 
    sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2) +
    sum((x11_seasadj[-1]-ssm_seasadj)^2)
    
  return(D)
}
```

Here we still let the variance to be *0.001, 0.01, 0.1, 1, 10, 100*.

```{r}  
# define the search function
exhaustion <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:6) {
     for (j in 1:6) {
       for (k in 1:6) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(10^(-4+j))) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 10^(-4+k)),
                   H = 10^(-4+i))
           ssm <- KFS(ssmm)
           
           dif <- Dif(x11, ssm, data)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, c(10^(-4+i),10^(-4+j),10^(-4+k)))
         }
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```{r}
unemp_exhaustion <- exhaustion(unemp)
unemp_exhaustion[which.min(unemp_exhaustion$difference),]
unemp_exhaustion
```

The result in the end is as same as what we got before, which is (10,10,10)

# Part 2

Again, we change the error into this defination: 
$$ \frac{\sum(X11_{irregular}-SSM_{irregular})^2}{\sigma_I^2} + \frac{\sum(X11_{trend}-SSM_{trend})^2}{\sigma_T^2} + \frac{\sum(X11_{seasonal}-SSM_{seasonal})^2}{\sigma_S^2}$$

```{r}
# change the defination of error  
Dif2 <- function(x11, ssm, data, sigma){
  
  x11_trend <- series(x11, 'd12')
  x11_seasonal <- series(x11, 'd10')
  x11_irregular <- series(x11, 'd13')
  
  ssm_trend <- coef(ssm, states = 'trend')
  ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  ssm_irregular <- data[-1] - ssm_trend[-1] - ssm_seasonal[-length(data)]
 
  D <-  sum((x11_irregular[-1]-ssm_irregular)^2)/sigma[1] + 
    sum((x11_trend-ssm_trend)^2)/sigma[2] + 
    sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2)/sigma[3] 
    
  return(D)
}
```


```{r}

# define the search function

exhaustion2 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:6) {
     for (j in 1:6) {
       for (k in 1:6) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(10^(-4+j))) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 10^(-4+k)),
                   H = 10^(-4+i))
           ssm <- KFS(ssmm)
           
           sigma <- c(10^(-4+i),10^(-4+j),10^(-4+k))
           
           dif <- Dif2(x11, ssm, data, sigma)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, sigma)
         }
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}

```

```{r}
unemp_exhaustion2 <- exhaustion2(unemp)
unemp_exhaustion2[which.min(unemp_exhaustion2$difference),]
```

Under this defination, our best fit is c(100,100,100). I basically have an idea why this happens, but let's see the result from this estimate first: 

```{r include=FALSE}

ssmm1 <- SSModel(unemp ~ SSMtrend(1, Q=list(10)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 10),
                   H = 10)
unemp_ssm1 <- KFS(ssmm1)

ssmm2 <- SSModel(unemp ~ SSMtrend(1, Q=list(100)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 100),
                   H = 100)
unemp_ssm2 <- KFS(ssmm2)

unemp_x11 <- seas(unemp, x11='')

unemp_x11_trend <- series(unemp_x11, 'd12')
unemp_x11_seasonal <- series(unemp_x11, 'd10')
unemp_x11_irregular <- series(unemp_x11, 'd13')

unemp_ssm1_trend <- coef(unemp_ssm1, states = 'trend')
unemp_ssm1_seasonal <- -rowSums(coef(unemp_ssm1, states='seasonal'))
unemp_ssm1_irregular <- unemp[-1] - unemp_ssm1_trend[-1] -
  unemp_ssm1_seasonal[-length(unemp)]

unemp_ssm2_trend <- coef(unemp_ssm2, states = 'trend')
unemp_ssm2_seasonal <- -rowSums(coef(unemp_ssm2, states='seasonal'))
unemp_ssm2_irregular <- unemp[-1] - unemp_ssm2_trend[-1] -
  unemp_ssm2_seasonal[-length(unemp)]
```

```{r echo=FALSE}
plot(unemp_x11_trend, ylim=c(5500,15500),ylab='')
par(new=TRUE)
plot(unemp_ssm2_trend, ylim=c(5500,15500),ylab='', col=2)
title(main='Trend') 
legend('topleft', c('x11','ssm'), col=c(1,2), lty=1)
```

```{r echo=FALSE}
plot(ts(unemp_x11_seasonal[-1], start=c(1990,2), frequency=12), 
     ylim=c(-700,1000),ylab='')
par(new=TRUE)
plot(ts(unemp_ssm2_seasonal[-length(unemp)],start=c(1990,2),frequency = 12), 
     col=2, ylim=c(-700,1000),ylab='')
title(main='Season(red:ssm)')
```

```{r echo=FALSE}
plot(ts(unemp_x11_seasonal[-1][1:100], start=c(1990,2), frequency=12), 
     ylim=c(-700,1000),ylab='')
par(new=TRUE)
plot(ts(unemp_ssm2_seasonal[-length(unemp)][1:100],start=c(1990,2),frequency = 12), 
     col=2, ylim=c(-700,1000),ylab='')
title(main='First 100 Season Points(red:ssm)')
```

```{r}
plot(ts(unemp_x11_irregular[-1], start=c(1990,02),frequency = 12), 
     ylim=c(-250,250), ylab='')
par(new=TRUE)
plot(ts(unemp_ssm2_irregular, start=c(1990,02),frequency = 12), 
     ylim=c(-250,250), ylab='', col=2)
title(main='Irregular(red:ssm)')
```

```{r}
plot(ts(unemp_x11_irregular[-1][1:100], start=c(1990,02),frequency = 12), 
     ylim=c(-250,250), ylab='')
par(new=TRUE)
plot(ts(unemp_ssm2_irregular[1:100], start=c(1990,02),frequency = 12), 
     ylim=c(-250,250), ylab='', col=2)
title(main='Irregular(red:ssm)')
```

```{r}
plot(unemp_ssm1_irregular, type='l')
par(new=TRUE)
plot(unemp_ssm2_irregular, type='l', col=2)
```

Based on the above figure, I **guess** that Kalman filter will give us the same result as long as those variances' values that we give it have the same ratio.


# Test my guess

```{r include=FALSE}

ssmm3 <- SSModel(unemp ~ SSMtrend(1, Q=list(1)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = 1)
unemp_ssm3 <- KFS(ssmm3)

unemp_ssm3_trend <- coef(unemp_ssm3, states = 'trend')
unemp_ssm3_seasonal <- -rowSums(coef(unemp_ssm3, states='seasonal'))
unemp_ssm3_irregular <- unemp[-1] - unemp_ssm3_trend[-1] -
  unemp_ssm3_seasonal[-length(unemp)]

```

```{r echo=FALSE}
# trend
plot(unemp_ssm1_trend, type='l')
par(new=TRUE)
plot(unemp_ssm2_trend, type='l', col=2)
par(new=TRUE)
plot(unemp_ssm3_trend, type='l', col=3)
```


```{r echo=FALSE}
# season
plot(unemp_ssm1_seasonal, type='l')
par(new=TRUE)
plot(unemp_ssm2_seasonal, type='l', col=2)
par(new=TRUE)
plot(unemp_ssm3_seasonal, type='l', col=3)
```

```{r echo=FALSE}
# irregular
plot(unemp_ssm1_irregular, type='l')
par(new=TRUE)
plot(unemp_ssm2_irregular, type='l', col=2)
par(new=TRUE)
plot(unemp_ssm3_irregular, type='l', col=3)
```

I have a further guess: as long as the ratio between these variances keep same, the results should be same.

# Test further guess

```{r include=FALSE}

ssmm4 <- SSModel(unemp ~ SSMtrend(1, Q=list(1)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 0.1),
                   H = 10)
unemp_ssm4 <- KFS(ssmm4)

unemp_ssm4_trend <- coef(unemp_ssm4, states = 'trend')
unemp_ssm4_seasonal <- -rowSums(coef(unemp_ssm4, states='seasonal'))
unemp_ssm4_irregular <- unemp[-1] - unemp_ssm4_trend[-1] -
  unemp_ssm4_seasonal[-length(unemp)]

ssmm5 <- SSModel(unemp ~ SSMtrend(1, Q=list(10)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = 100)
unemp_ssm5 <- KFS(ssmm5)

unemp_ssm5_trend <- coef(unemp_ssm5, states = 'trend')
unemp_ssm5_seasonal <- -rowSums(coef(unemp_ssm5, states='seasonal'))
unemp_ssm5_irregular <- unemp[-1] - unemp_ssm5_trend[-1] -
  unemp_ssm5_seasonal[-length(unemp)]
```

```{r echo=FALSE}
# trend
plot(unemp_ssm4_trend, type='l')
par(new=TRUE)
plot(unemp_ssm5_trend, type='l', col=2)
```

```{r echo=FALSE}
# season
plot(unemp_ssm4_seasonal, type='l')
par(new=TRUE)
plot(unemp_ssm5_seasonal, type='l', col=2)
```

```{r echo=FALSE}
# irregular
plot(unemp_ssm4_irregular, type='l')
par(new=TRUE)
plot(unemp_ssm5_irregular, type='l', col=2)
```

The results show that my guess is right.

**Comment: **these things should be figured out two months ago! So there is no reason to be happy for finding it!

**Note: ** cause the property I found above, I update the grid search choice then:

# version 2

## Use 'Dif' function defined at beginning 

Here, we take variances from 1:10 for each parameter:

```{r}
# define the search function
exhaustion_version2 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:10) {
     for (j in 1:10) {
       for (k in 1:10) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = k),
                   H = i)
           ssm <- KFS(ssmm)
           
           dif <- Dif(x11, ssm, data)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, c(i, j, k))
         }
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```{r}
unemp_exhaustionversion2 <- exhaustion_version2(unemp)
unemp_exhaustionversion2[which.min(unemp_exhaustionversion2$difference),]
```

If we enlarge the scale of parameters to 1:20:

```{r}
# define the search function
# this will cover the function above
exhaustion_version2 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:20) {
     for (j in 1:20) {
       for (k in 1:20) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = k),
                   H = i)
           ssm <- KFS(ssmm)
           
           dif <- Dif(x11, ssm, data)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, c(i, j, k))
         }
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```{r}
unemp_exhaustionversion2 <- exhaustion_version2(unemp)
unemp_exhaustionversion2[which.min(unemp_exhaustionversion2$difference),]
```


## Use 'Dif' function in Exhaustion2.Rmd(take product instead of sum)

```{r}
Dif2 <- function(x11, ssm, data){
  
  x11_trend <- series(x11, 'd12')
  x11_seasonal <- series(x11, 'd10')
  x11_seasadj <- series(x11, 'd11')
  
  ssm_trend <- coef(ssm, states = 'trend')
  ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  ssm_seasadj <- data[-1] - ssm_seasonal[-length(data)] # length is shorter
 
  D <-  sum((x11_trend-ssm_trend)^2) * 
    sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2) *
    sum((x11_seasadj[-1]-ssm_seasadj)^2)
  return(D)
}
```

Parameters take value in 1:10:

```{r}
# define the search function
exhaustion_version2_2 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:10) {
     for (j in 1:10) {
       for (k in 1:10) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = k),
                   H = i)
           ssm <- KFS(ssmm)
           
           dif <- Dif2(x11, ssm, data)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, c(i, j, k))
         }
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```{r}
unemp_exhaustionversion2_2 <- exhaustion_version2_2(unemp)
unemp_exhaustionversion2_2[which.min(unemp_exhaustionversion2_2$difference),]
```

Parameters take values in 1:20:

```{r}
# define the search function
# this will cover the function above
exhaustion_version2_2 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:20) {
     for (j in 1:20) {
       for (k in 1:20) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = k),
                   H = i)
           ssm <- KFS(ssmm)
           
           dif <- Dif2(x11, ssm, data)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, c(i, j, k))
         }
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```{r}
unemp_exhaustionversion2_2 <- exhaustion_version2_2(unemp)
unemp_exhaustionversion2_2[which.min(unemp_exhaustionversion2_2$difference),]
unemp_exhaustionversion2_2[17*20*20+11*20+7,]

```

**Based on the result above, we may take c(10,7,4) as the best value that we want.**