---
title: "Exhaustion&MLE&MAP"
author: "Linyi Guo"
date: "2019/11/2"
output: 
  html_document:
    toc: true
---

```{r include=FALSE}
rm(list=ls())
set.seed(9483)
```

```{r include=FALSE}
library(seasonal)
library(KFAS)
library(ggplot2)
library(plotly)
```

Based on **report20191030**, incorporated with Aaron's suggestion, make some modification of Exhaustion, MLE & MAP.

The related note is on my notebook, under topic **Review Meeting 10.30**.

# Exhaustion

Fix $\sigma_S^2=1$, we can achieve all possible combination under this condition, cause we care about the ratio among three variances instead of the value.

And we can use the second lose function now:
$\frac{\sum(X11_{irregular}-SSM_{irregular})^2}{\sigma_I^2} + \frac{\sum(X11_{trend}-SSM_{trend})^2}{\sigma_T^2} + \frac{\sum(X11_{seasonal}-SSM_{seasonal})^2}{\sigma_S^2}$

```{r}
# define the loss function
Dif <- function(x11, ssm, data, sigma){
  
  x11_trend <- series(x11, 'd12')
  x11_seasonal <- series(x11, 'd10')
  x11_irregular <- series(x11, 'd13')
  
  ssm_trend <- coef(ssm, states = 'trend')
  ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  ssm_irregular <- data[-1] - ssm_trend[-1] - ssm_seasonal[-length(data)]
 
  D <-  sum((x11_irregular[-1]-ssm_irregular)^2)/sigma[1] + 
    sum((x11_trend-ssm_trend)^2)/sigma[2] + 
    sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2)/sigma[3] 
    
  return(D)
}
```
```{r}

# define the search function

exhaustion <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:10) {
     for (j in 1:10) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j*0.2)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i*0.2)
           ssm <- KFS(ssmm)
           
           sigma <- c(i*0.2, j*0.2, 1)
           
           dif <- Dif(x11, ssm, data, sigma)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```
```{r}
unemp_exhaustion <- exhaustion(unemp)
unemp_exhaustion[which.min(unemp_exhaustion$difference),]

```

The above part is to illustrate that when fix $\sigma_S^2=1$, values of the other variances are greater than 1. That is to say we prefer to devide more variability to irregular and trend. **Therefore**, we can let them begin at 1, as following:


```{r}

# define the search function

exhaustion <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:20) {
     for (j in 1:20) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i)
           ssm <- KFS(ssmm)
           
           sigma <- c(i, j, 1)
           
           dif <- Dif(x11, ssm, data, sigma)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```
```{r}
unemp_exhaustion <- exhaustion(unemp)
unemp_exhaustion[which.min(unemp_exhaustion$difference),]
```

Thus the 'best' value(ratio) is c(7,5,1) when we set scope of variance is 1:20.(I tried scope 1:100 and the result is still c(7,5,1).)

Let's see the loss around this combination:

```{r echo=FALSE}
unemp_exhaustion[which(unemp_exhaustion$variance.1 %in% c(5,6,7,8,9) & 
                         unemp_exhaustion$variance.2 %in% c(3,4,5,6,7) ),]

difference <- matrix(unemp_exhaustion[which(unemp_exhaustion$variance.1 %in% c(5,6,7,8,9) &  unemp_exhaustion$variance.2 %in% c(3,4,5,6,7) ),]$difference,
                        nrow=5,byrow=TRUE)

plot_ly(showscale = FALSE) %>% add_surface(x=~c(3:7), y=~c(5:9), z=~difference)
```

And the results from c(7,5,1) are:

```{r include=FALSE}
ssmm <- SSModel(unemp ~ SSMtrend(1, Q=list(5)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = 7)
unemp_ssm <- KFS(ssmm)
  
unemp_x11 <- seas(unemp, x11='')

unemp_ssm_trend <- coef(unemp_ssm, states = 'trend')
unemp_ssm_seasonal <- -rowSums(coef(unemp_ssm, states='seasonal'))
unemp_ssm_irregular <- unemp[-1] - unemp_ssm_trend[-1] -
  unemp_ssm_seasonal[-length(unemp)]
unemp_ssm_seasadj <- unemp[-1]-unemp_ssm_seasonal[-length(unemp)]

unemp_x11_trend <- series(unemp_x11, 'd12')
unemp_x11_seasonal <- series(unemp_x11, 'd10')
unemp_x11_irregular <- series(unemp_x11, 'd13')
unemp_x11_seasadj <- series(unemp_x11, 'd11')
```

```{r echo=FALSE}
# Seasonal Adjustment
plot(ts(unemp_x11_seasadj[-1],start = c(1990,2),frequency = 12),
     ylim=c(5500,15500),ylab='')
par(new=TRUE)
plot(ts(unemp_ssm_seasadj, start = c(1990,2),frequency = 12), 
     ylim=c(5500,15500),ylab='', col=2)
title('Seasonal adjustment')
legend('topleft',c('X11','SSM'), col=c(1,2), lty=1)


# Trend
plot(unemp_x11_trend,ylim=c(5500,15500),ylab='')
par(new=TRUE)
plot(unemp_ssm_trend,ylim=c(5500,15500),ylab='', col=2)
title('Trend')
legend('topleft',c('X11','SSM'), col=c(1,2), lty=1)

# Season
plot(unemp_x11_seasonal[-1],type = 'l',ylim=c(-700,1000), ylab='')
par(new=TRUE)
plot(unemp_ssm_seasonal[-length(unemp)], type='l', ylim=c(-700,1000), ylab='',col=2)
title('Season')
legend('topleft',c('X11','SSM'), col=c(1,2), lty=1,cex = 0.8)


plot(unemp_x11_seasonal[-1][1:100],type = 'l',ylim=c(-700,1000), ylab='')
par(new=TRUE)
plot(unemp_ssm_seasonal[-length(unemp)][1:100], type='l', ylim=c(-700,1000), ylab='',col=2)
title('Season(the first 100 points)')
legend('topleft',c('X11','SSM'), col=c(1,2), lty=1,cex = 0.8)

# Irregular
plot(unemp_x11_irregular[-1], type='l', ylim=c(-450,500),ylab='')
par(new=TRUE)
plot(unemp_ssm_irregular, type='l', ylim=c(-450,500),ylab='', col=2)
title('Irregular')
legend('topleft',c('X11','SSM'), col=c(1,2), lty=1)
```

As a comparison, let's see the result if we let variance to be (1,1,1):

```{r include=FALSE}

ssmm <- SSModel(unemp ~ SSMtrend(1, Q=list(1)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = 1)
unemp_ssm111 <- KFS(ssmm)


unemp_ssm111_trend <- coef(unemp_ssm111, states = 'trend')
unemp_ssm111_seasonal <- -rowSums(coef(unemp_ssm111, states='seasonal'))
unemp_ssm111_irregular <- unemp[-1] - unemp_ssm111_trend[-1] -
  unemp_ssm111_seasonal[-length(unemp)]
unemp_ssm111_seasadj <- unemp[-1]-unemp_ssm111_seasonal[-length(unemp)]
```

```{r echo=FALSE}
# Seasonal Adjustment
plot(ts(unemp_x11_seasadj[-1],start = c(1990,2),frequency = 12),
     ylim=c(5500,15500),ylab='')
par(new=TRUE)
plot(ts(unemp_ssm111_seasadj, start = c(1990,2),frequency = 12), 
     ylim=c(5500,15500),ylab='', col=2)
par(new=TRUE)
plot(ts(unemp_ssm_seasadj, start = c(1990,2),frequency = 12), 
     ylim=c(5500,15500),ylab='', col=3)
title('Seasonal adjustment')
legend('topleft',c('X11','SSM111','SSM751'), col=c(1,2,3), lty=1)


# Trend
plot(unemp_x11_trend,ylim=c(5500,15500),ylab='')
par(new=TRUE)
plot(unemp_ssm111_trend,ylim=c(5500,15500),ylab='', col=2)
par(new=TRUE)
plot(unemp_ssm_trend,ylim=c(5500,15500),ylab='', col=3)
title('Trend')
legend('topleft',c('X11','SSM111','SSM751'), col=c(1,2,3), lty=1)

# Season
plot(unemp_x11_seasonal[-1],type = 'l',ylim=c(-700,1000), ylab='')
par(new=TRUE)
plot(unemp_ssm111_seasonal[-length(unemp)], type='l', ylim=c(-700,1000),
     ylab='',col=2)
par(new=TRUE)
plot(unemp_ssm_seasonal[-length(unemp)], type='l', ylim=c(-700,1000), 
     ylab='',col=3)
title('Season')
legend('topleft',c('X11','SSM111','SSM751'), col=c(1,2,3), lty=1,cex = 0.8)


plot(unemp_x11_seasonal[-1][1:100],type = 'l',ylim=c(-700,1000), ylab='')
par(new=TRUE)
plot(unemp_ssm111_seasonal[-length(unemp)][1:100], type='l', ylim=c(-700,1000),
     ylab='',col=2)
par(new=TRUE)
plot(unemp_ssm_seasonal[-length(unemp)][1:100], type='l', ylim=c(-700,1000),
     ylab='',col=3)
title('Season(the first 100 points)')
legend('topleft',c('X11','SSM111','SSM751'), col=c(1,2,3), lty=1,cex = 0.8)

# Irregular
plot(unemp_x11_irregular[-1], type='l', ylim=c(-450,500),ylab='')
par(new=TRUE)
plot(unemp_ssm111_irregular, type='l', ylim=c(-450,500),ylab='', col=2)
par(new=TRUE)
plot(unemp_ssm_irregular, type='l', ylim=c(-450,500),ylab='', col=3)
title('Irregular')
legend('topleft',c('X11','SSM111','SSM751'), col=c(1,2,3), lty=1)
```

From some details in the picture, it's generally true that the results from c(7,5,1) is better than those from c(1,1,1).

# MLE

```{r}
# define the log likelihood function
# because of seasonal components' characteristic
# I didn't compute the loglikelihood at first 11 points

loglikelihood <- function(data, trend, season, sigma){
  
  n <- length(data)
  a <- 0
  for (i in 12:n)  a <- a + (sum(season[(i-11):i]))^2
  l <- -(n-11)/2 * log(sigma[1]) - 
    (n-11)/2 * log(sigma[2]) - (n-11)/2 * log(sigma[3]) -
    sum((data[-c(1:11)]-trend[-c(1:11)]-season[-c(1:10,n)])^2)/(2*sigma[1]) - 
    sum((trend[-c(1:11)]-trend[-c(1:10,n)])^2)/(2*sigma[2]) - 
    a / (2*sigma[3])
  return(l)
  
}

```

Like the exhaustion part, we first check cases when variance of the other two is less than 1:

```{r}
# define the log likelihood matrix
loglikelihood_matrix <- function(data){
  LL <- c()
  index <- c()
   for (i in 1:100) {
     for (j in 1:100) {
         
         ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j*0.01)) + 
                   SSMseasonal(12, sea.type = 'dummy', Q = 1),
                 H = i*0.01)
         ssm <- KFS(ssmm)
         
         ssm_trend <- coef(ssm, states = 'trend')
         ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  #      ssm_seasadj <- data[-1] - ssm_seasonal[-length(data)] # length is shorter
         sigma <- c(i*0.01, j*0.01, 1)
         
         ll <- loglikelihood(data, ssm_trend, ssm_seasonal, sigma)
         LL <- c(LL, ll)
         index <- rbind(index, sigma)
      }
   }
  df <- data.frame(variance=index, loglikelihood=LL)
  return(df)
}
```

```{r}
unemp_loglikelihoodmatrix <- loglikelihood_matrix(unemp)
unemp_loglikelihoodmatrix[which.max(unemp_loglikelihoodmatrix$loglikelihood),]
```

Therefore, we may exclude the possibility that out mle is less than 1.

Let's look at the case when variances are greater than 1:

```{r}
# define the log likelihood matrix
loglikelihood_matrix <- function(data){
  LL <- c()
  index <- c()
   for (i in 1:100) {
     for (j in 1:100) {
         
         ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                   SSMseasonal(12, sea.type = 'dummy', Q = 1),
                 H = i)
         ssm <- KFS(ssmm)
         
         ssm_trend <- coef(ssm, states = 'trend')
         ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  #      ssm_seasadj <- data[-1] - ssm_seasonal[-length(data)] # length is shorter
         sigma <- c(i, j, 1)
         
         ll <- loglikelihood(data, ssm_trend, ssm_seasonal, sigma)
         LL <- c(LL, ll)
         index <- rbind(index, sigma)
      }
   }
  df <- data.frame(variance=index, loglikelihood=LL)
  return(df)
}
```

```{r}
unemp_loglikelihoodmatrix <- loglikelihood_matrix(unemp)
unemp_loglikelihoodmatrix[which.max(unemp_loglikelihoodmatrix$loglikelihood),]
```

After setting the scope at 1:25, 1:50 and 1:100, I find that the mle is always the largest value we could take for each scope and the ratio of max loglikelihood is approx 4:2:1. If we write down the expression of loglikelihood, we will found the variance is our denominator, so that is to say, our components' difference(numerator) doesn't have a huge change for different cases, and values of variances themselves play a more important role for loglikelihood.

Let's have a look at these loglikelihoods:

```{r echo=FALSE}
plot_ly(unemp_loglikelihoodmatrix,x=~variance.1, y=~variance.2, z=~loglikelihood,
        color = ~loglikelihood) %>% 
  add_markers()
```

# MAP






