---
title: "MLE&MAP3"
author: "LinyiGuo"
date: "2019/10/6"
output: 
  html_document:
    toc: true 
  pdf_document:
    toc: true
---

```{r echo = FALSE} 
rm(list = ls())
set.seed(9483)
```
```{r echo = FALSE}
library(KFAS)
library(tseries)
library(seasonal)
library(ggplot2)
```

# Grid search 

Grid search is to calculate the MLE or MAP, which could avoid local minimum value.

**Likelihood & Loglikelihood**

```{r}
# define likelihood function

likelihood <- function(theta, Trend, Season, Obs){
  n <- length(Obs)
  a <- 0 
  for (i in 12:n)  a <- a + (sum(Season[(i-11):i]))^2
  L <- (1/sqrt(2*pi*prod(theta)))^n * 
          exp(-sum((Obs - Trend - Season)^2)/(2*theta[1])) * 
          exp(-sum((Trend[-1]-Trend[-n])^2)/(2*theta[2])) *
          exp(- a/(2*theta[3]))
  return(L)
}


# loglikelihood
loglikelihood <- function(theta, Trend, Season, Obs ){
  n <- length(Obs)
  a <- 0
  for (i in 12:n)  a <- a + (sum(Season[(i-11):i]))^2
  l <- (-n/2) * log(theta[1]) - n/2 * log(theta[2]) - n/2 * log(theta[3]) -
    sum((Obs-Trend-Season)^2)/(2*theta[1]) - 
    sum((Trend[-1]-Trend[-n])^2)/(2*theta[2]) - 
    a / (2*theta[3])
  return(l)
}

```

## Grid Search for Single Parameter 

Here, I fix two variances at **1** and let the other one is a sequence with 10000 points.

1. variances in observation function are $0.001,0.002, \dots , 10$

2. variances in trend function are $0.01,0.02, \dots , 100$

3. variances in seasonal function are $0.01,0.02, \dots , 100$

### MLE

This is my code for related function:

```{r}
# define a 'large' function to return the loglikelihood series of one single parameter's sequence
loglikelihood_single_gs <- function(data, type){
  if(type == 1) {
    l1 <- c()
    for (i in 1:10000){
      ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=1), H=i*0.001)
      mod <- KFS(ssm)
      Trend <- coef(mod, states = 'trend')
      Season <- rowSums(coef(mod, states = 'seasonal'))
      theta <- c(i*0.001,1,1)
      l1 <- c(l1, loglikelihood(theta=theta, Trend=Trend, Season=Season, Obs = data))
    }
    return(l1)
  }
  
  if(type == 2){
    l2 <- c()
    for (i in 1:10000){
      ssm <- SSModel(data ~ SSMtrend(1, Q=list(i*0.01)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=1), H=1)
      mod <- KFS(ssm)
      Trend <- coef(mod, states = 'trend')
      Season <- rowSums(coef(mod, states = 'seasonal'))
      theta <- c(1,i*0.01,1)
      l2 <- c(l2, loglikelihood(theta=theta, Trend=Trend, Season=Season, Obs = data))
    }
    return(l2)
  }
  
  if(type == 3){
    l3 <- c()
    for (i in 1:10000){
      ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=i*0.01), H=1)
      mod <- KFS(ssm)
      Trend <- coef(mod, states = 'trend')
      Season <- rowSums(coef(mod, states = 'seasonal'))
      theta <- c(1,1,i*0.01)
      l3 <- c(l3, loglikelihood(theta=theta, Trend=Trend, Season=Season, Obs = data))
    }
    return(l3)
  }
}

```


**For the likelihood curve**

Here, the only difference from the function *loglikelihood_single_gs* is that we apply likelihood instead of loglikelihood defined at beginning.

```{r}
# define a 'large' function to return the loglikelihood series of one single parameter's sequence
likelihood_single_gs <- function(data, type){
  if(type == 1) {
    L1 <- c()
    for (i in 1:10000){
      ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=1), H=i*0.001)
      mod <- KFS(ssm)
      Trend <- coef(mod, states = 'trend')
      Season <- rowSums(coef(mod, states = 'seasonal'))
      theta <- c(i*0.001,1,1)
      L1 <- c(L1, likelihood(theta=theta, Trend=Trend, Season=Season, Obs = data))
    }
    return(L1)
  }
  
  if(type == 2){
    L2 <- c()
    for (i in 1:10000){
      ssm <- SSModel(data ~ SSMtrend(1, Q=list(i*0.01)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=1), H=1)
      mod <- KFS(ssm)
      Trend <- coef(mod, states = 'trend')
      Season <- rowSums(coef(mod, states = 'seasonal'))
      theta <- c(1,i*0.01,1)
      L2 <- c(L2, likelihood(theta=theta, Trend=Trend, Season=Season, Obs = data))
    }
    return(L2)
  }
  
  if(type == 3){
    L3 <- c()
    for (i in 1:10000){
      ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=i*0.01), H=1)
      mod <- KFS(ssm)
      Trend <- coef(mod, states = 'trend')
      Season <- rowSums(coef(mod, states = 'seasonal'))
      theta <- c(1,1,i*0.01)
      L3 <- c(L3, likelihood(theta=theta, Trend=Trend, Season=Season, Obs = data))
    }
    return(L3)
  }
}

```


The loglikelihood curves: 

```{r echo=FALSE}
data("AirPassengers")

output1_log <- loglikelihood_single_gs(data = log(AirPassengers), type = 1)

output2_log <- loglikelihood_single_gs(data = log(AirPassengers), type = 2)

output3_log <- loglikelihood_single_gs(data = log(AirPassengers), type = 3)
```
```{r echo=FALSE}
plot(output1_log, type = 'l')

plot(output2_log, type = 'l')

plot(output3_log, type = 'l')

```

The likelihood curves: 

```{r echo=FALSE}

output1_like <- likelihood_single_gs(data = log(AirPassengers), type = 1)

output2_like <- likelihood_single_gs(data = log(AirPassengers), type = 2)

output3_like <- likelihood_single_gs(data = log(AirPassengers), type = 3)

plot(output1_like, type = 'l')

plot(output2_like, type = 'l')

plot(output3_like, type = 'l')

```

I take a look at the first few points of case 1(var_y) specially since there is a change at the beginning:

```{r}
plot(output1_log[5:500], type = 'l', 
     main = '[5:500] points of Case Var_y', ylab = 'loglikelihood')

plot(output1_like[1:100], type = 'l', 
     main = 'first 100 points of Case Var_y', ylab = 'likelihood')


```




### MAP

Because our paramters are variances, that is to say they are greater than 0, so the priors on them should be some distribution that is defined on $(0,\infty)$.

**Prior**

```{r}
# if we let prior is standard exponential function
prior <- function(x) exp(-x)
```

**Posterior/MAP**

It is similar to the likelihood function but we times the prior:

```{r}
posterior_single_gs <- function(data, type) {
  
  if(type == 1){
    map1 <- c()
   for (i in 1:10000) {
     ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                      SSMseasonal(12, sea.type = 'dummy', Q = 1 ), H = i*0.001)
     mod <- KFS(ssm)
     Trend <- coef(mod, states = 'trend')
     Season <- rowSums(coef(mod, states = 'seasonal'))
     theta <- c(i*0.001,1,1)
     map <- likelihood(theta=theta,Trend=Trend,Season=Season,Obs = data) *
       prior(theta[type])
     map1 <- c(map1, map)
   } 
    return(map1)
  }
  
  if(type == 2){
    map2 <- c()
   for (i in 1:10000) {
     ssm <- SSModel(data ~ SSMtrend(1, Q=list(0.01*i)) + 
                      SSMseasonal(12, sea.type = 'dummy', Q = 1 ), H = 1)
     mod <- KFS(ssm)
     Trend <- coef(mod, states = 'trend')
     Season <- rowSums(coef(mod, states = 'seasonal'))
     theta <- c(1,i*0.01,1)
     map <- likelihood(theta=theta,Trend=Trend,Season=Season,Obs = data) *
       prior(theta[type])
     map2 <- c(map2, map)
   } 
    return(map2) 
  }
  
  if(type == 3){
    map3 <- c()
   for (i in 1:10000) {
     ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                      SSMseasonal(12, sea.type = 'dummy', Q = 0.01*i ), H = 1)
     mod <- KFS(ssm)
     Trend <- coef(mod, states = 'trend')
     Season <- rowSums(coef(mod, states = 'seasonal'))
     theta <- c(1,1,0.01*i)
     map <- likelihood(theta=theta,Trend=Trend,Season=Season,Obs = data) *
       prior(theta[type])
     map3 <- c(map3, map)
   } 
    return(map3)
  }
}

```

Then these are the posterior curves:(still on AirPassengers)

```{r echo=FALSE}
output1_post <- posterior_single_gs(data=log(AirPassengers), type = 1) 

output2_post <- posterior_single_gs(data=log(AirPassengers), type = 2) 

output3_post <- posterior_single_gs(data=log(AirPassengers), type = 3) 
```
```{r echo=FALSE}

ggplot(data.frame(output=output1_post, index = c(1:length(output1_post))),
       aes(x=index, y=output)) + 
  geom_line()

ggplot(data.frame(output=output2_post, index = c(1:length(output2_post))),
       aes(x=index, y=output)) + 
  geom_line()

ggplot(data.frame(output=output3_post, index = c(1:length(output3_post))),
       aes(x=index, y=output)) + 
  geom_line()


ggplot(data.frame(output=output2_post[5:100], 
                  index = c(5:100)),
       aes(x=index, y=output)) + 
  geom_line()
```


Let't take a look at the first 100 points of likelihood, prior and posterior curves at the same time(Fixed $\sigma_T^2, \sigma_S^2$ and $\sigma_y^2=0.001,0.002,\dots, 10$):

<center> **Orange: Likelihood ； Red: Prior  ；  Blue: Posterior** </center>
```{r echo=FALSE}
ggplot(data.frame(output1_like=output1_like[1:100],
                  output1_prior=prior(seq(0.001,10,0.001))[1:100],
                  output1_post=output1_post[1:100], 
                  index=c(1:100))) +
  geom_line(aes(y=output1_like, x=index),color='orange') + 
  geom_line(aes(y=output1_prior, x=index), color='red') +
  geom_line(aes(y=output1_post, x=index),color='blue') +
  labs(title = 'Case var_y', x='Index', y='') 

```

*Note: The prior curve is not a horizontal line actually.*

**If we change the prior to a generic exponential distribution**

```{r}
prior <- function(x, lambda) lambda * exp(-lambda * x)
```

**Update the posterior**(hide code)

```{r echo=FALSE}
posterior_single_gs <- function(data, type, lambda) { 
  # lambda is the parameter in our exponential prior function
  if(type == 1){
    map1 <- c()
   for (i in 1:10000) {
     ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                      SSMseasonal(12, sea.type = 'dummy', Q = 1 ), H = i*0.001)
     mod <- KFS(ssm)
     Trend <- coef(mod, states = 'trend')
     Season <- rowSums(coef(mod, states = 'seasonal'))
     theta <- c(i*0.001,1,1)
     map <- likelihood(theta=theta,Trend=Trend,Season=Season,Obs = data) *
       prior(theta[type],lambda)
     map1 <- c(map1, map)
   } 
    return(map1)
  }
  
  if(type == 2){
    map2 <- c()
   for (i in 1:10000) {
     ssm <- SSModel(data ~ SSMtrend(1, Q=list(0.01*i)) + 
                      SSMseasonal(12, sea.type = 'dummy', Q = 1 ), H = 1)
     mod <- KFS(ssm)
     Trend <- coef(mod, states = 'trend')
     Season <- rowSums(coef(mod, states = 'seasonal'))
     theta <- c(1,i*0.01,1)
     map <- likelihood(theta=theta,Trend=Trend,Season=Season,Obs = data) *
       prior(theta[type],lambda)
     map2 <- c(map2, map)
   } 
    return(map2) 
  }
  
  if(type == 3){
    map3 <- c()
   for (i in 1:10000) {
     ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                      SSMseasonal(12, sea.type = 'dummy', Q = 0.01*i ), H = 1)
     mod <- KFS(ssm)
     Trend <- coef(mod, states = 'trend')
     Season <- rowSums(coef(mod, states = 'seasonal'))
     theta <- c(1,1,0.01*i)
     map <- likelihood(theta=theta,Trend=Trend,Season=Season,Obs = data) *
       prior(theta[type],lambda)
     map3 <- c(map3, map)
   } 
    return(map3)
  }
}
```

We let lambda = 100:

```{r echo=FALSE}
lambda <- 100

output1_post <- posterior_single_gs(data=log(AirPassengers), type = 1, lambda) 

output2_post <- posterior_single_gs(data=log(AirPassengers), type = 2, lambda) 

output3_post <- posterior_single_gs(data=log(AirPassengers), type = 3, lambda) 
```
```{r echo=FALSE}

ggplot(data.frame(output=output1_post, index = c(1:length(output1_post))),
       aes(x=index, y=output)) + 
  geom_line()

ggplot(data.frame(output=output2_post, index = c(1:length(output2_post))),
       aes(x=index, y=output)) + 
  geom_line()

ggplot(data.frame(output=output3_post, index = c(1:length(output3_post))),
       aes(x=index, y=output)) + 
  geom_line()
```

```{r}
# var_y
ggplot(data.frame(output1_like=output1_like[1:100],
                  output1_prior=prior(seq(0.001,10,0.001),lambda)[1:100],
                  output1_post=output1_post[1:100], 
                  index=c(1:100))) +
  geom_line(aes(y=output1_like, x=index),color='orange') + 
  geom_line(aes(y=output1_prior, x=index), color='red') +
  geom_line(aes(y=output1_post, x=index),color='blue') +
  labs(title = 'Case Var_y', x='Index', y='') 

plot(output1_post[1:100], type = 'l')
which.max(output1_post)
```

```{r}
# var_T

ggplot(data.frame(output2_like=output2_like[1:10],
                  output2_prior=prior(seq(0.01,100,0.01),lambda)[1:10],
                  output2_post=output2_post[1:10], 
                  index=c(1:10))) +
  geom_line(aes(y=output2_like, x=index),color='orange') + 
  geom_line(aes(y=output2_prior, x=index), color='red') +
  geom_line(aes(y=output2_post, x=index),color='blue') +
  labs(title = 'Case Var_T', x='Index', y='') 


```

```{r}
# var_S
ggplot(data.frame(output3_like=output3_like[1:10],
                  output3_prior=prior(seq(0.01,100,0.01),lambda)[1:10],
                  output3_post=output3_post[1:10], 
                  index=c(1:10))) +
  geom_line(aes(y=output3_like, x=index),color='orange') + 
  geom_line(aes(y=output3_prior, x=index), color='red') +
  geom_line(aes(y=output3_post, x=index),color='blue') +
  labs(title = 'Case Var_S', x='Index', y='') 
```


<center> **Again, Orange: Likelihood ； Red: Prior  ；  Blue: Posterior** </center>

# Compare with the results from TRAMO-SEATS

The model here is too bad, cause what I did is just choosing three map values from three cases(fix two find one).  
<center>**Just take it for fun** </center>

```{r echo=FALSE}
library(seasonal)
library(KFAS)
  mod_x11 <- seas(AirPassengers, x11='')
Trend_x11 <- series(mod_x11, 'd12')
Season_x11 <- series(mod_x11, 'd10')

mod_ssm <- SSModel(log(AirPassengers) ~ SSMtrend(1, Q=list(0.01)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q=0.01), H = 36*0.001)
mod_ssm <- KFS(mod_ssm)
Trend_ssm <- exp(coef(mod_ssm, states='trend'))
Season_ssm <- ts(exp(-rowSums(coef(mod_ssm, states='seasonal'))),
                 start = c(1949,1), frequency = 12)
```

```{r}
plot(Trend_ssm, ylab='',ylim=c(0,600))
par(new=TRUE)
plot(Trend_x11, ylab='',ylim=c(0,600), col=2)
legend('topleft',c('ssm','x11'), col=c(1,2), lty=c(1,1))
```

**NOTE/UPDATE: I just found my code to compute the seasonal component in ssm of this document is wrong. I forgot to add minus sign in the expression. And this mistake could happen in the last two mle/map files too... (2019.10.22)**

```{r}
plot(Season_ssm[-length(AirPassengers)], ylab='',type = 'l', ylim=c(0.75,1.25))
par(new=TRUE)
plot(Season_x11[-1], ylab= '', col=2, ylim=c(0.75,1.25), lwd=2, type='l')
```


# Improvement on MLE&MAP

Before we move on, draw some figures to show that our likelihood curve/posterior curve is decreasing all the time:

## case var_T

### likelihood
```{r}
ggplot(data = data.frame(output = output2_like[1:10], index=c(1:10)),
       aes(y=output, x=index)) + 
  geom_line()

ggplot(data = data.frame(output = output2_like[10:100], index=c(10:100)),
       aes(y=output, x=index)) + 
  geom_line()

ggplot(data = data.frame(output = output2_like[50:100], index=c(50:100)),
       aes(y=output, x=index)) + 
  geom_line()

ggplot(data = data.frame(output = output2_like[100:200], index=c(100:200)),
       aes(y=output, x=index)) + 
  geom_line()


ggplot(data = data.frame(output = output2_like[1000:2000], index=c(1000:2000)),
       aes(y=output, x=index)) + 
  geom_line()

```

Actually, as long as the difference between each two points is negative, we can say that likelihood decreases all the time.

```{r}

dif2_like <- output2_like[-1] - output2_like[-10000]

dif2_like <- ifelse(dif2_like > 0, 1, 0)

sum(dif2_like)
```

Similarly, check the var_S:
```{r echo=FALSE}
dif3_like <- output3_like[-1] - output3_like[-10000]

dif3_like <- ifelse(dif3_like > 0, 1, 0)

sum(dif3_like)
```

### posterior

The difference is defined to 0 if actual difference is negative; 1 if actual difference is positve. Just like what we defined before.

Case 2:

```{r echo=FALSE}

dif2_post <- output2_post[-1] - output2_post[-10000]

dif2_post <- ifelse(dif2_post > 0, 1, 0)

sum(dif2_post)
```

Case 3:

```{r echo=FALSE}

dif3_post <- output3_post[-1] - output3_post[-10000]

dif3_post <- ifelse(dif3_post > 0, 1, 0)

sum(dif3_post)
```


## Improvement on MLE

```{r improvement}

likelihood_single_gs <- function(data, type, gap1, gap2, gap3){
  if(type == 1) {
    L1 <- c()
    for (i in 1:10000){
      ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=1), H=i*gap1)
      mod <- KFS(ssm)
      Trend <- coef(mod, states = 'trend')
      Season <- rowSums(coef(mod, states = 'seasonal'))
      theta <- c(i*gap1,1,1)
      L1 <- c(L1, likelihood(theta=theta, Trend=Trend, Season=Season, Obs = data))
    }
    return(L1)
  }
  
  if(type == 2){
    L2 <- c()
    for (i in 1:10000){
      ssm <- SSModel(data ~ SSMtrend(1, Q=list(i*gap2)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=1), H=1)
      mod <- KFS(ssm)
      Trend <- coef(mod, states = 'trend')
      Season <- rowSums(coef(mod, states = 'seasonal'))
      theta <- c(1,i*gap2,1)
      L2 <- c(L2, likelihood(theta=theta, Trend=Trend, Season=Season, Obs = data))
    }
    return(L2)
  }
  
  if(type == 3){
    L3 <- c()
    for (i in 1:10000){
      ssm <- SSModel(data ~ SSMtrend(1, Q=list(1)) + 
                       SSMseasonal(12, sea.type = 'dummy', Q=i*gap3), H=1)
      mod <- KFS(ssm)
      Trend <- coef(mod, states = 'trend')
      Season <- rowSums(coef(mod, states = 'seasonal'))
      theta <- c(1,1,i*gap3)
      L3 <- c(L3, likelihood(theta=theta, Trend=Trend, Season=Season, Obs = data))
    }
    return(L3)
  }
}
```

# Switch to Exhaustion 1
 