---
title: "post_normal"
author: "Linyi Guo"
date: "2019/11/25"
output: 
  html_document:
    toc: true
---
```{r include=FALSE}
rm(list=ls())
set.seed(9483)
```

```{r include=FALSE}
library(seasonal)
library(KFAS)
library(ggplot2)
library(plotly)
```

```{r include=FALSE}
Data1 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data1.csv')
Data2 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data2.csv')
Data3 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data3.csv')
Data4 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data4.csv')
Data5 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data5.csv')
Data6 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data6.csv')
Data7 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data7.csv')
Data8 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data8.csv')
Data9 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data9.csv')
Data0 <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\simulation&reproduction\\data\\data0.csv')
```
```{r}
data1 <- ts(Data1[,2], start=c(2000,01), frequency=12)
data2 <- ts(Data2[,2], start=c(2000,01), frequency=12)
data3 <- ts(Data3[,2], start=c(2000,01), frequency=12)
data4 <- ts(Data4[,2], start=c(2000,01), frequency=12)
data5 <- ts(Data5[,2], start=c(2000,01), frequency=12)
data6 <- ts(Data6[,2], start=c(2000,01), frequency=12)
data7 <- ts(Data7[,2], start=c(2000,01), frequency=12)
data8 <- ts(Data8[,2], start=c(2000,01), frequency=12)
data9 <- ts(Data9[,2], start=c(2000,01), frequency=12)
data0 <- ts(Data0[,2], start=c(2000,01), frequency=12)



data1_x11 <- seas(data1, x11='')
data2_x11 <- seas(data2, x11='')
data3_x11 <- seas(data3, x11='')
data4_x11 <- seas(data4, x11='')
data5_x11 <- seas(data5, x11='')
data6_x11 <- seas(data6, x11='')
data7_x11 <- seas(data7, x11='')
data8_x11 <- seas(data8, x11='')
data9_x11 <- seas(data9, x11='')
data0_x11 <- seas(data0, x11='')
```

```{r}
Datalist <- list(data1=data1, data2=data2, data3=data3, data4=data4, data5=data5, 
                 data6=data6, data7=data7, data8=data8, data9=data9, data0=data0)
```

```{r echo=FALSE}
par(mfrow=c(2,2))
plot(data1_x11)
plot(series(data1_x11, 'd13'))
plot(series(data1_x11, 'd12'))
plot(series(data1_x11, 'd10'))

plot(data2_x11)
plot(series(data2_x11, 'd13'))
plot(series(data2_x11, 'd12'))
plot(series(data2_x11, 'd10'))

plot(data3_x11)
plot(series(data3_x11, 'd13'))
plot(series(data3_x11, 'd12'))
plot(series(data3_x11, 'd10'))

plot(data4_x11)
plot(series(data4_x11, 'd13'))
plot(series(data4_x11, 'd12'))
plot(series(data4_x11, 'd10'))

plot(data5_x11)
plot(series(data5_x11, 'd13'))
plot(series(data5_x11, 'd12'))
plot(series(data5_x11, 'd10'))

plot(data6_x11)
plot(series(data6_x11, 'd13'))
plot(series(data6_x11, 'd12'))
plot(series(data6_x11, 'd10'))

plot(data7_x11)
plot(series(data7_x11, 'd13'))
plot(series(data7_x11, 'd12'))
plot(series(data7_x11, 'd10'))

plot(data8_x11)
plot(series(data8_x11, 'd13'))
plot(series(data8_x11, 'd12'))
plot(series(data8_x11, 'd10'))

plot(data9_x11)
plot(series(data9_x11, 'd13'))
plot(series(data9_x11, 'd12'))
plot(series(data9_x11, 'd10'))

plot(data0_x11)
plot(series(data0_x11, 'd13'))
plot(series(data0_x11, 'd12'))
plot(series(data0_x11, 'd10'))
par(mfrow=c(1,1))
```

**These work is not hard but very time-consuming.**

# Bunch of functions


```{r}
# loss
Dif1 <- function(x11, ssm, data, sigma){
  
  x11_trend <- series(x11, 'd12')
  x11_seasonal <- series(x11, 'd10')
  x11_irregular <- series(x11, 'd13')
  
  ssm_trend <- coef(ssm, states = 'trend')
  ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  ssm_irregular <- data[-1] - ssm_trend[-1] - ssm_seasonal[-length(data)]
 
  D <-  sum((x11_irregular[-1]-ssm_irregular)^2)/sigma[1] + 
    sum((x11_trend-ssm_trend)^2)/sigma[2] + 
    sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2)/sigma[3] 
    
  return(D)
}



# exhaustion

exhaustion1 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:20) {
     for (j in 1:20) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i)
           ssm <- KFS(ssmm)
           
           sigma <- c(i, j, 1)
           
           dif <- Dif1(x11, ssm, data, sigma)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}

# loglikelihood

loglikelihood <- function(data, trend, season, sigma){
  
  n <- length(data)
  a <- 0
  for (i in 12:n)  a <- a + (sum(season[(i-11):i]))^2
  l <- -(n-11)/2 * log(sigma[1]) - 
    (n-11)/2 * log(sigma[2]) - (n-11)/2 * log(sigma[3]) -
    sum((data[-c(1:11)]-trend[-c(1:11)]-season[-c(1:10,n)])^2)/(2*sigma[1]) - 
    sum((trend[-c(1:11)]-trend[-c(1:10,n)])^2)/(2*sigma[2]) - 
    a / (2*sigma[3])
  return(l)
  
}

# logprior


logprior2 <- function(sigma,c1,c2,p){
  return(-c1*p*sigma[1]- c2*p*sigma[2])
}

# logposterior

logposterior2_matrix <- function(data, c1, c2, p){
  LP <- c()
  index <- c()
   for (i in 1:15) {
     for (j in 1:15) {
         
         ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                   SSMseasonal(12, sea.type = 'dummy', Q = 1),
                 H = i)
         ssm <- KFS(ssmm)
         
         ssm_trend <- coef(ssm, states = 'trend')
         ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  #      ssm_seasadj <- data[-1] - ssm_seasonal[-length(data)] # length is shorter
         sigma <- c(i, j, 1)
         
         lp <- loglikelihood(data, ssm_trend, ssm_seasonal, sigma) + logprior2(sigma, c1, c2,p)
         LP <- c(LP, lp)
         index <- rbind(index, sigma)
         
      }
   }
  df <- data.frame(variance=index, logposterior=LP)
  return(df)
}
```



```{r}

error <- function(data,i,j,N) {
  data <- ts(as.data.frame(data),start = c(2000,1),frequency = 12)

  x11 <- seas(data,x11='')
  data_ssmmN <- SSModel(data~SSMtrend(1,Q=list(N)) + 
                          SSMseasonal(12,sea.type = 'dummy', Q=1), H=N)
  data_ssmN <- KFS(data_ssmmN)
  data_ssmN_trend <- coef(data_ssmN, states = 'trend')
  data_ssmN_seasonal <- -rowSums(coef(data_ssmN, states = 'seasonal'))

  
  p <- -loglikelihood(data,data_ssmN_trend,data_ssmN_seasonal,c(N,N,1))/N
  
  data_postmatrix <- logposterior2_matrix(data,i,j,p)
  map <- as.numeric(data_postmatrix[which.max(data_postmatrix$logposterior),][-4])
                                 
      
  ssmm <- SSModel(data ~ SSMtrend(1, Q=list(map[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map[1])
  ssm <- KFS(ssmm)
  error <- Dif1(x11,ssm,data,map)
    
  return(error) 
}


Prior2_gridsearch <- function(datalist) {
  N <- 20
  Error <- c()
  index <- c()
  
  for (i in 1:1) {
    for (j in 1:1) {
      errorlist <- sapply(datalist, error)
      error <- sum(errorlist)
      Error <- c(Error,error)
      index <- rbind(index, c(i,j))
      
    }
    
  }
  df <- data.frame(c1=index[,1],c2=index[,2],Error=Error)
  return(df)
}

Prior2_gridsearch(Datalist)
```

```
c1c2_gridsearch <- Prior2_gridsearch(Datalist)

write.csv(c1c2_gridsearch,'C:\\Users\\GuoLY\\Desktop\\markdown\\MLE & MAP\\prior2_c1c2_simulation.csv')
```

It needs a lot of time to run so I saved the final resul here. **UPDATE: the result is wrong here!**

```{r}
c1c2_gridsearch[which.min(c1c2_gridsearch$Error),]

```

Well the final result is **c(1,1)** !!!???

Let's see the exhaustion results of these 10 datasets:

```{r}
exhaustionlist <- lapply(Datalist, exhaustion1)
```


```{r}
Exhaustion <- rbind(
exhaustionlist$data1[which.min(exhaustionlist$data1$difference),],
exhaustionlist$data2[which.min(exhaustionlist$data2$difference),],
exhaustionlist$data3[which.min(exhaustionlist$data3$difference),],
exhaustionlist$data4[which.min(exhaustionlist$data4$difference),],
exhaustionlist$data5[which.min(exhaustionlist$data5$difference),],
exhaustionlist$data6[which.min(exhaustionlist$data6$difference),],
exhaustionlist$data7[which.min(exhaustionlist$data7$difference),],
exhaustionlist$data8[which.min(exhaustionlist$data8$difference),],
exhaustionlist$data9[which.min(exhaustionlist$data9$difference),],
exhaustionlist$data0[which.min(exhaustionlist$data0$difference),]
)

rownames(Exhaustion) <- c()
Exhaustion
```

If we use the c1=c2=1 then the map estimators are :

```{r}
plist <- sapply(Datalist, function(data,N=20){
  
  data <- ts(as.data.frame(data),start = c(2000,1),frequency = 12)

  x11 <- seas(data,x11='')
  data_ssmmN <- SSModel(data~SSMtrend(1,Q=list(N)) + 
                          SSMseasonal(12,sea.type = 'dummy', Q=1), H=N)
  data_ssmN <- KFS(data_ssmmN)
  data_ssmN_trend <- coef(data_ssmN, states = 'trend')
  data_ssmN_seasonal <- -rowSums(coef(data_ssmN, states = 'seasonal'))

  
  p <- -loglikelihood(data,data_ssmN_trend,data_ssmN_seasonal,c(N,N,1))/N
  
  return(p)
  
})


plist
data1_postmatrix <- logposterior2_matrix(data1,1,1,plist[1])
data1_postmatrix[which.max(data1_postmatrix$logposterior),]
```



I am so frustrated now. There is always one bug and the worse thing is I can't fix it. Keep tring! You can do it!!!


I decide to use the stupid method for now:

```{r}
Prior2_gridsearch <- function(datalist){
  Error <- c()
  index <- c()
  x11list <- lapply(datalist, function(data) seas(data,x11='') )
  
  
  for (i in 1:10) {
    for (j in 1:10) {
      
      ############# 1
      
      data1_postmatrix <- logposterior2_matrix(data1,i,j,plist[1])
      map1<-as.numeric(data1_postmatrix[which.max(data1_postmatrix$logposterior),][-4])
      
      ssmm1 <- SSModel(data1 ~ SSMtrend(1, Q=list(map1[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map1[1])
      ssm1 <- KFS(ssmm1)
      dif1 <- Dif1(x11list$data1,ssm1,data1,map1)
      
      
      
      ############## 2
      
      data2_postmatrix <- logposterior2_matrix(data2,i,j,plist[2])
      map2<-as.numeric(data2_postmatrix[which.max(data2_postmatrix$logposterior),][-4])
      
      ssmm2 <- SSModel(data2 ~ SSMtrend(1, Q=list(map2[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map2[1])
      ssm2 <- KFS(ssmm2)
      dif2 <- Dif1(x11list$data2,ssm2,data2,map2)
      
      
      ############## 3
      
      data3_postmatrix <- logposterior2_matrix(data3,i,j,plist[3])
      map3<-as.numeric(data3_postmatrix[which.max(data3_postmatrix$logposterior),][-4])
      
      ssmm3 <- SSModel(data3 ~ SSMtrend(1, Q=list(map3[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map3[1])
      ssm3 <- KFS(ssmm3)
      dif3 <- Dif1(x11list$data3,ssm3,data3,map3)
      
      
      ############## 4
      
      data4_postmatrix <- logposterior2_matrix(data4,i,j,plist[4])
      map4<-as.numeric(data4_postmatrix[which.max(data4_postmatrix$logposterior),][-4])
      
      ssmm4 <- SSModel(data4 ~ SSMtrend(1, Q=list(map4[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map4[1])
      ssm4 <- KFS(ssmm4)
      dif4 <- Dif1(x11list$data4,ssm4,data4,map4)
      
      
      ############## 5
      
      data5_postmatrix <- logposterior2_matrix(data5,i,j,plist[5])
      map5<-as.numeric(data5_postmatrix[which.max(data5_postmatrix$logposterior),][-4])
      
      ssmm5 <- SSModel(data5 ~ SSMtrend(1, Q=list(map5[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map5[1])
      ssm5 <- KFS(ssmm5)
      dif5 <- Dif1(x11list$data5,ssm5,data5,map5)
      
      
      ############## 6
      
      data6_postmatrix <- logposterior2_matrix(data6,i,j,plist[6])
      map6<-as.numeric(data6_postmatrix[which.max(data6_postmatrix$logposterior),][-4])
      
      ssmm6 <- SSModel(data6 ~ SSMtrend(1, Q=list(map6[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map6[1])
      ssm6 <- KFS(ssmm6)
      dif6 <- Dif1(x11list$data6,ssm6,data6,map6)
      
      
      ############## 7 
      
      data7_postmatrix <- logposterior2_matrix(data7,i,j,plist[7])
      map7<-as.numeric(data7_postmatrix[which.max(data7_postmatrix$logposterior),][-4])
      
      ssmm7 <- SSModel(data7 ~ SSMtrend(1, Q=list(map7[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map7[1])
      ssm7 <- KFS(ssmm7)
      dif7 <- Dif1(x11list$data7,ssm7,data7,map7)
      
      
      ############## 8 
      
      data8_postmatrix <- logposterior2_matrix(data8,i,j,plist[8])
      map8<-as.numeric(data8_postmatrix[which.max(data8_postmatrix$logposterior),][-4])
      
      ssmm8 <- SSModel(data8 ~ SSMtrend(1, Q=list(map8[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map8[1])
      ssm8 <- KFS(ssmm8)
      dif8 <- Dif1(x11list$data8,ssm8,data8,map8)
      
      
      ############## 9
      
      data9_postmatrix <- logposterior2_matrix(data9,i,j,plist[9])
      map9<-as.numeric(data9_postmatrix[which.max(data9_postmatrix$logposterior),][-4])
      
      ssmm9 <- SSModel(data1 ~ SSMtrend(1, Q=list(map9[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map9[1])
      ssm9 <- KFS(ssmm9)
      dif9 <- Dif1(x11list$data9,ssm9,data9,map9)
      
      
      ############## 0
      
      data0_postmatrix <- logposterior2_matrix(data0,i,j,plist[10])
      map0<-as.numeric(data0_postmatrix[which.max(data0_postmatrix$logposterior),][-4])
      
      ssmm0 <- SSModel(data0 ~ SSMtrend(1, Q=list(map0[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = map0[1])
      ssm0 <- KFS(ssmm0)
      dif0 <- Dif1(x11list$data0,ssm0,data0,map0)
      
      
      
      error <- dif1+dif2+dif3+dif4+dif5+dif6+dif7+dif8+dif9+dif0
      Error <- c(Error, error)
      index <- rbind(index,c(i,j))
    }
    
  }
  df <- data.frame(c1=index[,1],c2=index[,2],Error=Error)
  return(df)
}
```

```
c1c2_gridsearch <- Prior2_gridsearch(Datalist)


write.csv(c1c2_gridsearch,'C:\\Users\\GuoLY\\Desktop\\markdown\\MLE & MAP\\prior2_c1c2_simulation.csv')
```

well, it is obvious that we need to use apply() in the future cause programming this stupid method is too time-consuming. 

```{r}
c1c2_gridsearch[which.min(c1c2_gridsearch$Error),]
```

Well it's still 11................................

```{r}


error <- function(data,i,j,p) {
  
  x11 <- seas(data, x11='')

  data_postmatrix <- logposterior2_matrix(data,i,j,p)
  
  map <- as.numeric(data_postmatrix[which.max(data_postmatrix$logposterior),][-4])
                                 
      
  ssmm <- SSModel(data ~ SSMtrend(1, Q=list(map[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1), H = map[1])
  
  ssm <- KFS(ssmm)
  
  error <- Dif1(x11,ssm,data,map)
    
  return(error) 
}

Prior2_gridsearch <- function(datalist) {
  Error <- c()
  index <- c()
  
  for (i in 1:3) {
    for (j in 1:3) {
      error1 <- error(datalist$data1,i,j,plist[1])
      error2 <- error(datalist$data2,i,j,plist[2])
      error3 <- error(datalist$data3,i,j,plist[3])
      error4 <- error(datalist$data4,i,j,plist[4])
      error5 <- error(datalist$data5,i,j,plist[5])
      error6 <- error(datalist$data6,i,j,plist[6])
      error7 <- error(datalist$data7,i,j,plist[7])
      error8 <- error(datalist$data8,i,j,plist[8])
      error9 <- error(datalist$data9,i,j,plist[9])
      error0 <- error(datalist$data0,i,j,plist[10])
      
      error <- error1+error2+error3+error4+error5+error6+error7+error8+error9+error0
      
      Error <- c(Error, error)
      index <- rbind(index, c(i,j))
      
    }
    
  }
  
  df <- data.frame(c1=index[,1],c2=index[,2],Error=Error)
  return(df)
}

```

```{r}
Prior2_gridsearch(Datalist)
```

试了一个3*3的小loop发现我之前那个结果应该是错了。

```{r}


error <- function(data,i,j,p) {
  
  x11 <- seas(data, x11='')

  data_postmatrix <- logposterior2_matrix(data,i,j,p)
  
  map <- as.numeric(data_postmatrix[which.max(data_postmatrix$logposterior),][-4])
                                 
      
  ssmm <- SSModel(data ~ SSMtrend(1, Q=list(map[2])) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1), H = map[1])
  
  ssm <- KFS(ssmm)
  
  error <- Dif1(x11,ssm,data,map)
    
  return(error) 
}

Prior2_gridsearch <- function(datalist) {
  Error <- c()
  index <- c()
  
  for (i in 1:10) {
    for (j in 1:10) {
      error1 <- error(datalist$data1,i,j,plist[1])
      error2 <- error(datalist$data2,i,j,plist[2])
      error3 <- error(datalist$data3,i,j,plist[3])
      error4 <- error(datalist$data4,i,j,plist[4])
      error5 <- error(datalist$data5,i,j,plist[5])
      error6 <- error(datalist$data6,i,j,plist[6])
      error7 <- error(datalist$data7,i,j,plist[7])
      error8 <- error(datalist$data8,i,j,plist[8])
      error9 <- error(datalist$data9,i,j,plist[9])
      error0 <- error(datalist$data0,i,j,plist[10])
      
      error <- error1+error2+error3+error4+error5+error6+error7+error8+error9+error0
      
      Error <- c(Error, error)
      index <- rbind(index, c(i,j))
      
    }
    
  }
  
  df <- data.frame(c1=index[,1],c2=index[,2],Error=Error)
  return(df)
}

```

```
c1c2_gridsearch <- Prior2_gridsearch(Datalist)

write.csv(c1c2_gridsearch,'C:\\Users\\GuoLY\\Desktop\\markdown\\MLE & MAP\\prior2_c1c2_simulation.csv')
```

```{r}
c1c2_gridsearch[which.min(c1c2_gridsearch$Error),]
```

Although this result is still not good, let's try it for a moment:

```{r}
data1_postmatrix <- logposterior2_matrix(data1,1,8,plist[1])
map1 <- as.numeric(data1_postmatrix[which.max(data1_postmatrix$logposterior),])

data2_postmatrix <- logposterior2_matrix(data2,1,8,plist[2])
map2 <- as.numeric(data2_postmatrix[which.max(data2_postmatrix$logposterior),])

data3_postmatrix <- logposterior2_matrix(data3,1,8,plist[3])
map3 <- as.numeric(data3_postmatrix[which.max(data3_postmatrix$logposterior),])

data4_postmatrix <- logposterior2_matrix(data4,1,8,plist[4])
map4 <- as.numeric(data4_postmatrix[which.max(data4_postmatrix$logposterior),])

data5_postmatrix <- logposterior2_matrix(data5,1,8,plist[5])
map5 <- as.numeric(data5_postmatrix[which.max(data5_postmatrix$logposterior),])

data6_postmatrix <- logposterior2_matrix(data6,1,8,plist[6])
map6 <- as.numeric(data6_postmatrix[which.max(data6_postmatrix$logposterior),])

data7_postmatrix <- logposterior2_matrix(data7,1,8,plist[7])
map7 <- as.numeric(data7_postmatrix[which.max(data7_postmatrix$logposterior),])

data8_postmatrix <- logposterior2_matrix(data8,1,8,plist[8])
map8 <- as.numeric(data8_postmatrix[which.max(data8_postmatrix$logposterior),])

data9_postmatrix <- logposterior2_matrix(data9,1,8,plist[9])
map9 <- as.numeric(data9_postmatrix[which.max(data9_postmatrix$logposterior),])

data0_postmatrix <- logposterior2_matrix(data0,1,8,plist[10])
map0 <- as.numeric(data0_postmatrix[which.max(data0_postmatrix$logposterior),])


Map <- rbind(map1,map2,map3,map4,map5,map6,map7,map8,map9,map0)

Map
```

Well the results are not good. I am quite wondering why this happens. Let's check whether the likelihood behave like before:

```{r}
# loglikelihood

loglikelihood <- function(data, trend, season, sigma){
  
  n <- length(data)
  a <- 0
  for (i in 12:n)  a <- a + (sum(season[(i-11):i]))^2
  l <- -(n-11)/2 * log(sigma[1]) - 
    (n-11)/2 * log(sigma[2]) - (n-11)/2 * log(sigma[3]) -
    sum((data[-c(1:11)]-trend[-c(1:11)]-season[-c(1:10,n)])^2)/(2*sigma[1]) - 
    sum((trend[-c(1:11)]-trend[-c(1:10,n)])^2)/(2*sigma[2]) - 
    a / (2*sigma[3])
  return(l)
  
}

# define the loglikelihood matrix
loglikelihood_matrix <- function(data){
  LL <- c()
  index <- c()
   for (i in 1:50) {
     for (j in 1:50) {
         
         ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                   SSMseasonal(12, sea.type = 'dummy', Q = 1),
                 H = i)
         ssm <- KFS(ssmm)
         
         ssm_trend <- coef(ssm, states = 'trend')
         ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  #      ssm_seasadj <- data[-1] - ssm_seasonal[-length(data)] # length is shorter
         sigma <- c(i, j, 1)
         
         ll <- loglikelihood(data, ssm_trend, ssm_seasonal, sigma)
         LL <- c(LL, ll)
         index <- rbind(index, sigma)
      }
   }
  df <- data.frame(variance=index, loglikelihood=LL)
  return(df)
}
```


```{r}
data1_loglikelihoodmatrix <- loglikelihood_matrix(data1)
data1_loglikelihoodmatrix[which.max(data1_loglikelihoodmatrix$loglikelihood),]
```

Well the likelihood behaves like what we saw before in the file *Posterior1*.

We may find that the map estimators are quite large so I enlarge the **p**s' values as following(from N=20 to N=15):

```{r}

exhaustion1 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
   for (i in 1:15) {
     for (j in 1:15) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i)
           ssm <- KFS(ssmm)
           
           sigma <- c(i, j, 1)
           
           dif <- Dif1(x11, ssm, data, sigma)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}

exhaustionlist <- lapply(Datalist, exhaustion1)


Exhaustion2 <- rbind(
exhaustionlist$data1[which.min(exhaustionlist$data1$difference),],
exhaustionlist$data2[which.min(exhaustionlist$data2$difference),],
exhaustionlist$data3[which.min(exhaustionlist$data3$difference),],
exhaustionlist$data4[which.min(exhaustionlist$data4$difference),],
exhaustionlist$data5[which.min(exhaustionlist$data5$difference),],
exhaustionlist$data6[which.min(exhaustionlist$data6$difference),],
exhaustionlist$data7[which.min(exhaustionlist$data7$difference),],
exhaustionlist$data8[which.min(exhaustionlist$data8$difference),],
exhaustionlist$data9[which.min(exhaustionlist$data9$difference),],
exhaustionlist$data0[which.min(exhaustionlist$data0$difference),]
)

rownames(Exhaustion2) <- c()
Exhaustion2
```


```{r}
plist <- sapply(Datalist, function(data,N=15){
  
  data <- ts(as.data.frame(data),start = c(2000,1),frequency = 12)

  x11 <- seas(data,x11='')
  data_ssmmN <- SSModel(data~SSMtrend(1,Q=list(N)) + 
                          SSMseasonal(12,sea.type = 'dummy', Q=1), H=N)
  data_ssmN <- KFS(data_ssmmN)
  data_ssmN_trend <- coef(data_ssmN, states = 'trend')
  data_ssmN_seasonal <- -rowSums(coef(data_ssmN, states = 'seasonal'))

  
  p <- -loglikelihood(data,data_ssmN_trend,data_ssmN_seasonal,c(N,N,1))/N
  
  return(p)
  
})
```
```
c1c2_gridsearch <- Prior2_gridsearch(Datalist)

write.csv(c1c2_gridsearch,'C:\\Users\\GuoLY\\Desktop\\markdown\\MLE & MAP\\prior2_c1c2_simulationN15.csv')

```

```{r}
c1c2_gridsearch[which.min(c1c2_gridsearch$Error),]

```


```{r}
data1_postmatrix <- logposterior2_matrix(data1,2,10,plist[1])
map1 <- as.numeric(data1_postmatrix[which.max(data1_postmatrix$logposterior),])

data2_postmatrix <- logposterior2_matrix(data2,2,10,plist[2])
map2 <- as.numeric(data2_postmatrix[which.max(data2_postmatrix$logposterior),])

data3_postmatrix <- logposterior2_matrix(data3,2,10,plist[3])
map3 <- as.numeric(data3_postmatrix[which.max(data3_postmatrix$logposterior),])

data4_postmatrix <- logposterior2_matrix(data4,2,10,plist[4])
map4 <- as.numeric(data4_postmatrix[which.max(data4_postmatrix$logposterior),])

data5_postmatrix <- logposterior2_matrix(data5,2,10,plist[5])
map5 <- as.numeric(data5_postmatrix[which.max(data5_postmatrix$logposterior),])

data6_postmatrix <- logposterior2_matrix(data6,2,10,plist[6])
map6 <- as.numeric(data6_postmatrix[which.max(data6_postmatrix$logposterior),])

data7_postmatrix <- logposterior2_matrix(data7,2,10,plist[7])
map7 <- as.numeric(data7_postmatrix[which.max(data7_postmatrix$logposterior),])

data8_postmatrix <- logposterior2_matrix(data8,2,10,plist[8])
map8 <- as.numeric(data8_postmatrix[which.max(data8_postmatrix$logposterior),])

data9_postmatrix <- logposterior2_matrix(data9,2,10,plist[9])
map9 <- as.numeric(data9_postmatrix[which.max(data9_postmatrix$logposterior),])

data0_postmatrix <- logposterior2_matrix(data0,2,10,plist[10])
map0 <- as.numeric(data0_postmatrix[which.max(data0_postmatrix$logposterior),])


Map2 <- rbind(map1,map2,map3,map4,map5,map6,map7,map8,map9,map0)

Map2
```














