---
title: "Reproduction of StatCan"
author: "LinyiGuo"
date: "2019/8/20"
output:
  html_document: default
  pdf_document: default
---
```{r}
rm(list = ls())
options(digits = 4)
set.seed(9483)
```

# Import data and load packages

```{r}
library(ggfortify)
library(forecast)
library(KFAS)
library(tseries)
library(gridExtra)
library(seasonal)

# Load Data
data_nike <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\nike.csv',header = TRUE)
data_adi <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\adidas.csv',header = TRUE)

data_nike <- ts(rev(data_nike[3:58,2]), frequency=4, start=c(2005,1))
data_adi <- ts(rev(data_adi[2:57,2]), frequency=4, start=c(2005,1))
data("AirPassengers")
data_ap <- AirPassengers

# visualize data
plot(data_nike,ylab='',main="Nike Quarterly Sales from 2005 to 2018")
plot(data_adi,ylab='',main="Adidas Quarterly Sales from 2005 to 2018")
plot(data_ap,ylab='',main="Air Passengers from 1949 to 1960")
plot(log(data_ap),ylab='',main="Log of Air Passengers from 1949 to 1960")

```


# Construct Models

```{r}
# nike
## build TRAMO-SEATS
mod_nike_seats <- seas(data_nike)

## build x-11
mod_nike_x11 <- seas(data_nike, x11='')

## build ssm
mod_nike_ssm <- SSModel(data_nike ~ SSMtrend(1, Q = list(1)) + 
                        SSMseasonal(4,Q=0.1, sea.type = "trigonometric"), H=1)
out_nike_ssm <- KFS(mod_nike_ssm)


# adi
## build TRAMO-SEATS
mod_adi_seats <- seas(data_adi)

## build x-11
mod_adi_x11 <- seas(data_adi, x11='')

## build ssm
mod_adi_ssm <- SSModel(data_adi ~ SSMtrend(1, Q = list(1)) + 
                        SSMseasonal(4,Q=0.1, sea.type = "trigonometric"), H=1)
out_adi_ssm <- KFS(mod_adi_ssm)


# airpassengers
## build TRAMO-SEATS
mod_ap_seats <- seas(data_ap)

## build x-11
mod_ap_x11 <- seas(data_ap, x11='')

## build ssm
mod_ap_ssm <- SSModel(log(data_ap) ~ SSMtrend(1, Q = list(1)) + 
                        SSMseasonal(12,Q=0.1, sea.type = "trigonometric"), H=1)
out_ap_ssm <- KFS(mod_ap_ssm)

```

# Seasonal adjustment 

```{r echo=FALSE}
# nike
plot(series(mod_nike_seats,"s11"),ylab='',ylim=c(3000,10000))
par(new=TRUE)
plot(series(mod_nike_x11,"d11"),ylab='',ylim=c(3000,10000),col=2,lty=2)
par(new=TRUE)
plot(data_nike-rowSums(coef(out_nike_ssm,states = 'seasonal')[,c(1,3)]),
     ylab='',ylim=c(3000,10000),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Seasonal adjusted results of Nike quarterly sales data from 2005 to 2018')

# adi
plot(series(mod_adi_seats,"s11"),ylab='',ylim=c(1000,7000))
par(new=TRUE)
plot(series(mod_adi_x11,"d11"),ylab='',ylim=c(1000,7000),col=2,lty=2)
par(new=TRUE)
plot(data_adi-rowSums(coef(out_adi_ssm,states = 'seasonal')[,c(1,3)]),
     ylab='',ylim=c(1000,7000),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Seasonal adjusted results of Adidas quarterly sales data from 2005 to 2018')

# airpassengers
plot(series(mod_ap_seats,"s11"),ylab='',ylim=c(100,500))
par(new=TRUE)
plot(series(mod_ap_x11,"d11"),ylab='',ylim=c(100,500),col=2,lty=2)
par(new=TRUE)
plot(exp(log(data_ap)-
       rowSums(coef(out_ap_ssm,states = 'seasonal')[,c(1,3,5,7,9,11)])),
     ylab='',ylim=c(100,500),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Seasonal adjusted results of air passengers from 1949 to 1960')
```

# Trend

```{r echo=FALSE}
# nike
plot(series(mod_nike_seats,"s12"),ylab='',ylim=c(3000,10000))
par(new=TRUE)
plot(series(mod_nike_x11,"d12"),ylab='',ylim=c(3000,10000),col=2,lty=2)
par(new=TRUE)
plot(coef(out_nike_ssm,states = 'trend'),
     ylab='',ylim=c(3000,10000),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Trend component of Nike quarterly sales data from 2005 to 2018')

# adi
plot(series(mod_adi_seats,"s12"),ylab='',ylim=c(1000,7000))
par(new=TRUE)
plot(series(mod_adi_x11,"d12"),ylab='',ylim=c(1000,7000),col=2,lty=2)
par(new=TRUE)
plot(coef(out_adi_ssm,states = 'trend'),
     ylab='',ylim=c(1000,7000),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Trend component of Adidas quarterly sales data from 2005 to 2018')

# airpassengers
plot(series(mod_ap_seats,"s12"),ylab='',ylim=c(100,500))
par(new=TRUE)
plot(series(mod_ap_x11,"d12"),ylab='',ylim=c(100,500),col=2,lty=2)
par(new=TRUE)
plot(coef(out_ap_ssm,states = 'trend'),
     ylab='',ylim=c(100,500),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Trend component of air passengers from 1949 to 1960')

```

# Seasonal

```{r echo=FALSE}
# nike
plot(series(mod_nike_seats,"s10"),ylab='',ylim=c(-400,500))
par(new=TRUE)
plot(series(mod_nike_x11,"d10"),ylab='',ylim=c(-400,500),col=2,lty=2)
par(new=TRUE)
plot(ts(rowSums(coef(out_nike_ssm,states = 'seasonal')[,c(1,3)]),
        frequency = 4, start = c(2005,1)),
     ylab='',ylim=c(-400,500),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Seasonal influence of Nike quarterly sales data from 2005 to 2018')

# adi
plot(series(mod_adi_seats,"s10"),ylab='',ylim=c(-700,700))
par(new=TRUE)
plot(series(mod_adi_x11,"d10"),ylab='',ylim=c(-700,700),col=2,lty=2)
par(new=TRUE)
plot(ts(rowSums(coef(out_adi_ssm,states = 'seasonal')[,c(1,3)]),
        frequency = 4, start = c(2005,1)),
     ylab='',ylim=c(-700,700),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Seasonal influence of Adidas quarterly sales data from 2005 to 2018')

# airpassengers
plot(series(mod_ap_seats,"s10"),ylab='',ylim=c(0.7,1.4))
par(new=TRUE)
plot(series(mod_ap_x11,"d10"),ylab='',ylim=c(0.7,1.4),col=2,lty=2)
par(new=TRUE)
plot(ts(exp(rowSums(coef(out_ap_ssm,states = 'seasonal')[,c(1,3,5,7,9,11)])),
        frequency = 12, start = c(1949,1)),
     ylab='',ylim=c(0.7,1.4),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Seasonal influence of air passengers from 2005 to 2018')

```

# Residuals

```{r echo=FALSE}
# nike
plot(series(mod_nike_seats,"s13"),ylab='',ylim=c(-650,300))
par(new=TRUE)
plot(series(mod_nike_x11,"d13"),ylab='',ylim=c(-650,300), col=2,lty=2)
par(new=TRUE)
plot(residuals(out_nike_ssm,"response"),
     ylab='',ylim=c(-650,300),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Residuals of Nike quarterly sales data from 2005 to 2018')

# adi
plot(series(mod_adi_seats,"s13"),ylab='',ylim=c(-200,180))
par(new=TRUE)
plot(series(mod_adi_x11,"d13"),ylab='',ylim=c(-200,180), col=2,lty=2)
par(new=TRUE)
plot(residuals(out_adi_ssm,"response"),
     ylab='',ylim=c(-200,180),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Residuals of Adidas quarterly sales data from 2005 to 2018')

# air passengers
plot(series(mod_ap_seats,"s13"),ylab='',ylim=c(0.94,1.11))
par(new=TRUE)
plot(series(mod_ap_x11,"d13"),ylab='',ylim=c(0.94,1.11),col=2,lty=2)
par(new=TRUE)
plot(exp(residuals(out_ap_ssm,"response")),
     ylab='',ylim=c(0.94,1.11),col=3,lty=4)
legend('topleft',c('seats','x11','ssm'),col=c(1,2,3),lty=c(1,2,4))
title('Residuals of air passengers from 1949 to 1960')

```

# Different Situations

## Small residuals

The residuals of ssm from statcan is almost invariant so we can realize it by forcing `H` in `SSModel` to be 0(or some very very small number, like 0.0001).

```{r}
mod_nike_nonoi <- SSModel(data_nike ~ SSMtrend(1, Q = list(1)) + 
                            SSMseasonal(4, Q = 0.1, "trigonometric"), 
                          H = 0)
out_nike_nonoi <- KFS(mod_nike_nonoi)

# plot residuals
plot(series(mod_nike_seats,"s13"),ylab='',ylim=c(-650,300))
par(new=TRUE)
plot(series(mod_nike_x11,"d13"),ylab='',ylim=c(-650,300), col=2,lty=2)
par(new=TRUE)
plot(residuals(out_nike_nonoi,"response"),
     ylab='',ylim=c(-650,300),col=3,lty=4)
legend('topleft',c('seats','x11','ssm(nonoise)'),col=c(1,2,3),lty=c(1,2,4))
title('Residuals of Nike quarterly sales data from 2005 to 2018')


# check the trend and seasonal
## trend
plot(series(mod_nike_seats,"s12"),ylab='',ylim=c(3000,10000))
par(new=TRUE)
plot(series(mod_nike_x11,"d12"),ylab='',ylim=c(3000,10000), col=2,lty=2)
par(new=TRUE)
plot(coef(out_nike_nonoi,states="trend"),
     ylab='',ylim=c(3000,10000),col=3,lty=4)
legend('topleft',c('seats','x11','ssm(nonoise)'),col=c(1,2,3),lty=c(1,2,4))
title('Trends of Nike quarterly sales data from 2005 to 2018')

## seasonal
plot(series(mod_nike_seats,"s10"),ylab='',ylim=c(-400,500))
par(new=TRUE)
plot(series(mod_nike_x11,"d10"),ylab='',ylim=c(-400,500), col=2,lty=2)
par(new=TRUE)
plot(ts(rowSums(coef(out_nike_nonoi,states = 'seasonal')[,c(1,3)]),
        frequency = 4, start = c(2005,1)),
     ylab='',ylim=c(-400,500),col=3,lty=4)
legend('topleft',c('seats','x11','ssm(nonoise)'),col=c(1,2,3),lty=c(1,2,4))
title('Seasonal component of Nike quarterly sales data from 2005 to 2018')

```
```{r include=FALSE}
rowSums(coef(out_nike_nonoi)[,c(1,2,4)]) - data_nike
rowSums(coef(out_nike_ssm)[,c(1,2,4)])+residuals(out_nike_ssm)-data_nike
rowSums(coef(out_ap_ssm)[,c(1,2,4,6,8,10,12)])+residuals(out_ap_ssm)-data_ap

rowSums(series(mod_nike_seats,c('s10','s12','s13'))) - data_nike
rowSums(series(mod_adi_seats,c('s10','s12','s13'))) - data_adi
plot(rowSums(series(mod_ap_seats,c('s10','s12','s13'))) - data_ap)

residuals(out_nike_ssm,"recursive")
residuals(out_nike_nonoi,"recursive")
summary(mod_nike_seats)
summary(mod_nike_x11)
summary(seas(data_nike,outlier = NULL))
summary(seas(data_nike,x11='',outlier = NULL))
summary(mod_ap_seats)

```


## Plot the proportion of each component for the whole the movement

The expression from statcan to calculate the proportion is $$SEAS=\frac{\sum_{t=2}^N((y_t-S_t)-(y_{t-1}-S_{t-1}))^2}{\sum_{t=2}^N(y_t-y_{t-1})^2}$$ And the absorption of trend and irregular is calculated by further removing the appropriate component from each term.

which means the proportion of trend will be $$TREND=\frac{\sum_{t=2}^N((y_t-S_t-T_t)-(y_{t-1}-S_{t-1}-T_{t-1}))^2}{\sum_{t=2}^N(y_t-y_{t-1})^2}$$

Personally I don't think it's correct. 

**So we need another method to measure this kind proportion**






