---
title: "BuildThePrior"
author: "Linyi Guo"
date: "2020/1/8"
output: 
  html_document:
    toc: true
---


```{r include=FALSE}
set.seed(9483)
rm(list=ls())
```

```{r}
# install these packages if needed
#install.packages("seasonal")
#install.packages("forecast")
#install.packages("KFAS")

# import package
library(seasonal)
library(forecast)
library(KFAS)
```

# SIMULATION

To avoid potential mistakes, I only simulate the datasets from ARIMA(1,1,1)(0,1,1)


## Define a function to simulate data that we want

```{r}
simulation1 <- function(length){
     
   model <- Arima(ts(rnorm(120),start=c(1980,01),frequency =12), order=c(1,1,1),   
                  seasonal=c(0,1,1), fixed=c(phi=runif(1), theta=runif(1),
                                             Theta=runif(1))
                  )
   
   data <- simulate(model, nsim=length)
   
   # because if we need to take log later, data must be positive
   if(min(data) <= 0) data <- data - min(data) + runif(1)
   else data <- data
   
   return(data)
  
}
```


## function to simulate a datalist with a lot of datasets

```{r}
simlist1 <- function(n,length) {   
  
  Datalist <- list()
  
  for (i in 1:n)  Datalist[[i]] <- simulation1(length)
  
  return(Datalist)  
  
}

```

```{r}
set.seed(9483)

datalist2 <- simlist1(100, 180)

# If time permits, you can try this one
# datalist2 <- simlist1(200, 180)
 
```


Until this step, datalist is basically what I want to use in the following analysis, but we still have one more step: that is to preprocessing.




# Preprocessing

```{r}
# define a function for outliers and log-transformation
preprocess <- function(x11) {
  
  if(transformfunction(x11) == 'log') 
    data <- log(series(x11, 'b1'))
  else
    data <- series(x11, 'b1')

  return(data)
}
```


```{r}
# build model list
x11list2 <- lapply(datalist2, function(x) seas(x, x11=''))

```


```{r}
# obtain the preprocessed datalist denoted as Datalist
Datalist2 <- lapply(x11list2, preprocess)

```

Now we finished this step, then we need to find the 'ideal' values of parameters in these datasets.


# Building Our Prior

## Searching for the 'best' value

**Loss function**

```{r}

Dif1 <- function(x11, ssm, data, sigma){
  
  x11_trend <- series(x11, 'd12')
  x11_seasonal <- series(x11, 'd10')
  x11_irregular <- series(x11, 'd13')
  
  ssm_trend <- coef(ssm, states = 'trend')
  ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  ssm_irregular <- data[-1] - ssm_trend[-1] - ssm_seasonal[-length(data)]
 
  D <-  sum((x11_irregular[-1]-ssm_irregular)^2)/sigma[1] + 
    sum((x11_trend-ssm_trend)^2)/sigma[2] + 
    sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2)/sigma[3] 
    
  return(D)
}
```

**Exhaustion function**

```{r}

exhaustion1 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
  
   for (i in 1:100) {
     for (j in 1:100) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j*0.2)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i*0.2)
           ssm <- KFS(ssmm)
           
           sigma <- c(i*0.2, j*0.2, 1)
           
           dif <- Dif1(x11, ssm, data, sigma)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```{r}
# system.time(exhaustion1(Datalist1[[1]]))
#   user  system  elapsed
#  189.80 237.43  427.96 

idevallist2 <- lapply(Datalist2, exhaustion1)
```

```{r}
idevalmat2 <- c()

for (i in 1:100){
  ideval <- idevallist2[[i]][which.min(idevallist2[[i]]$difference),c(1,2)]
  idevalmat2 <- rbind(idevalmat2, ideval)
}

```

And then please output the data frame `idevalmat2`.

```
write.csv(idevalmat2, "The position//idevalmat2.csv")
```

