---
title: "airline model(Kalman filter)"
author: "LinyiGuo"
date: "2019/8/8"
output:
  pdf_document: default
  html_document: default
---

# Import Data

```{r}
library(KFAS)
library(ggfortify)
library(forecast)
data("AirPassengers")
data_ap_log <- log(AirPassengers)
autoplot(AirPassengers)+labs(x = "year", y = "# of Passengers", 
                             title = "Monthly Air Passengers from 1949 to 1961")
autoplot(data_ap_log)+labs(x = "year", y = "log of Passengers", 
                             title = "log of Monthly Air Passengers from 1949 to 1961")
```

**Again**, since the difference is larger as time goes on, we take log first.

# Results from ARIMA

```{r}
decompose_ap1 <- decompose(data_ap_log, "additive")
autoplot(decompose_ap1)
```


# Build SSModel and Apply Kalman Filter

According to [paper 3.5](https://www.jstatsoft.org/article/view/v039i02/v39i02.pdf), **KFAS** follows closely Durbin and Koopman's book and papers.

```{r}
# create one SSModel object
mod_ap <- SSModel(data_ap_log~ SSMtrend(1, Q=list(1)) + 
                    SSMseasonal(period = 12, sea.type = "dummy"),
                    H = 1)  # dummy here means we don't use trigonometric expression
print(mod_ap) 

# apply kalman filter 
out_ap <- KFS(mod_ap, filtering = "state")
print(out_ap)

# get estimates of conditional means of states
coef(out_ap, states = "trend")
head(coef(out_ap, states = "seasonal"))

# compare seasonal components
par(mfcol=c(2,1))
plot.ts(coef(out_ap, states = "seasonal")[,1], ylab = "", 
        main = "kalman filter $ season")
plot.ts(decompose_ap1$seasonal, ylab = '', main = "Arima$season")
par(mfcol=c(1,1))
```

## Use trigonometric expression 

**HINT**: the seasonal component is the sum of column 1,3,5,7,9,11, when we use trigonometric seasonal.

```{r}
mod_ap_tri <- SSModel(data_ap_log ~ SSMtrend(1, Q=list(5)) + 
                    SSMseasonal(period = 12, sea.type = "trigonometric"),
                  H = 1) 
out_ap_tri <- KFS(mod_ap_tri, filtering = "state")
print(out_ap_tri)


par(mfcol=c(3,1))
plot.ts(coef(out_ap_tri, states = "seasonal")[,1], ylab="")
plot.ts(ts(rowSums(coef(mod_ap_tri, states = "seasonal")[,c(1,3,5,7,9,11)]),
                   frequency = 12, start = 1949)) # sum the col's to get trigonometric season
plot.ts(decompose_ap1$seasonal)
par(mfcol=c(1,1))

plot.ts(coef(out_ap_tri, states = "seasonal")[,-1]) # maximum is 10 ts
```

### compare the seasonal components from Arima and Kalman

```{r}
plot.ts(ts(rowSums(coef(mod_ap_tri, states = "seasonal")[,c(1,3,5,7,9,11)]),
           frequency = 12, start = 1949), col="red", lwd = 1.5, 
        ylim=c(-0.3,0.3), ylab='')
par(new=TRUE)
plot.ts(ts(decompose_ap1$seasonal,frequency = 12, start = 1949),
        lty = 2, ylim=c(-0.3,0.3),ylab='')
title(main = 'seasonal component')
legend("topleft",c("Kalman", "Arima"),col=c("red","black"), lty=c(1,2),cex=0.8)
```

### compare the trend components from Arima and Kalman

It shows that the *trend* from Kalman filter includes noise.

```{r}
plot.ts(ts(coef(mod_ap_tri, states = "trend"),
           frequency = 12, start = 1949), col="red", lwd = 1.5, 
       ylim=c(4.5,6.5), ylab='')
par(new=TRUE)
plot.ts(ts(decompose_ap1$trend,frequency = 12, start = 1949),
        lty = 2, ylim=c(4.5,6.5),ylab='') # only trend from arima
title(main = 'trend component')
legend("topleft",c("Kalman", "Arima"),col=c("red","black"), lty=c(1,2),cex=0.8)

```

Two curves fit almost perfectly after adding noise to arima's trend.

```{r}
plot.ts(ts(coef(mod_ap_tri, states = "trend"),
           frequency = 12, start = 1949), col="red", lwd = 1.5, 
        ylim=c(4.5,6.5), ylab='')
par(new=TRUE)
plot.ts(ts(decompose_ap1$trend+decompose_ap1$random,frequency = 12, start = 1949),
        lty = 2, ylim=c(4.5,6.5),ylab='') # sum trend and noise from arima
title(main = 'trend component')
legend("topleft",c("Kalman", "Arima"),col=c("red","black"), lty=c(1,2),cex=0.8)
```

# Seasonal Adjustment

Since the results of seasonal components are the same, the deseasoned seires should be the same as well.

## arima
```{r}
autoplot(data_ap_log-decompose_ap1$seasonal)+labs(x="date", y="log of deseasoned series", title = " log of original series after deseasoning (Arima)") 



plot.ts(exp(data_ap_log-decompose_ap1$seasonal),
        ylim=c(80,620),ylab='',xlab='', col="red",lwd=1.5)
par(new=TRUE)
plot.ts(AirPassengers,ylim=c(80,620),ylab='',xlab='')
title(xlab = "date", ylab="deseasoned series", main = "original series after deseasoning (Arima)") 

```

## kalman filter

```{r}
autoplot(data_ap_log - rowSums(coef(mod_ap_tri, 
                                    states = "seasonal")[,c(1,3,5,7,9,11)])) + labs(x="date", y="log of deseasoned series", title = " log of original series after deseasoning (Kalman Filter)")

plot.ts(exp(data_ap_log - rowSums(coef(mod_ap_tri, 
                                    states = "seasonal")[,c(1,3,5,7,9,11)])),
        xlab='', ylab='', ylim=c(80,620), col='red', lwd=1.5)
par(new=TRUE)
plot.ts(AirPassengers,xlab='', ylab='', ylim=c(80,620))
title(xlab="date", ylab="deseasoned series", main = "original series after deseasoning (Kalman Filter)")
```


# Some improvement

**the variance of trend is 0.5**

```{r}
mod_ap_tri2 <- SSModel(data_ap_log ~ SSMtrend(1, Q=list(0.5)) + 
                    SSMseasonal(period = 12, sea.type = "trigonometric"),
                  H = 1) 
out_ap_tri2 <- KFS(mod_ap_tri2, filtering = "state")
print(out_ap_tri2)

plot(coef(out_ap_tri2, states = "trend"), ylim=c(4.5,6.5),ylab='',col=2)
par(new=TRUE)
plot(decompose_ap1$trend,ylim=c(4.5,6.5),ylab='', lty=2)
title(main='Trend comparison')
legend('topleft', c('Kalman(0.5)', 'Arima'), col=c(2,1), lty = c(1,2), cex = 0.8)

```

**the variance of trend is 0.1**

```{r}
mod_ap_tri3 <- SSModel(data_ap_log ~ SSMtrend(1, Q=list(0.1)) + 
                    SSMseasonal(period = 12, sea.type = "trigonometric"),
                  H = 1) 
out_ap_tri3 <- KFS(mod_ap_tri3, filtering = "state")
print(out_ap_tri3)


# compare trend
plot(coef(out_ap_tri3, states = "trend"), ylim=c(4.5,6.5),ylab='',col=2)
par(new=TRUE)
plot(decompose_ap1$trend,ylim=c(4.5,6.5),ylab='', lty=2)
title(main='trend comparison')
legend('topleft', c('Kalman(0.1)', 'Arima'), col=c(2,1), lty = c(1,2), cex = 0.8)


# let's compare seasonal part under this case
plot(ts(rowSums(coef(out_ap_tri3, states = "seasonal")[,c(1,3,5,7,9,11)]),
        frequency = 12, start = 1949),
     ylim=c(-.3,.3),ylab='',col=2)
par(new=TRUE)
plot(decompose_ap1$seasonal,ylim=c(-.3,.3),ylab='', lty=2)
title(main =' seasonal component ' )
legend('topleft', c('Kalman(0.1)', 'Arima'), col=c(2,1), lty = c(1,2), cex = 0.8)

```

Until here, we still don't reproduce the results from statcan, where the seasonal adjustment curve is smooth.(**SOLVED IN KALMAN FILTER DOCUMENT**)