---
title: "airline model"
author: "linyiguo"
date: "2019/7/25"
output:
  html_document: default
  pdf_document: default
---

## Data Exploratory
```{r}
library(forecast)
library(tseries)
library(ggfortify)
data("AirPassengers")
data_ap <- AirPassengers

data_ap
sum(is.na(data_ap))
frequency(data_ap)
cycle(data_ap)
summary(data_ap)
plot(data_ap, xlab="Date", ylab="Passenger numbers(1000's)", 
     main = "Air passenger numbers from 1949 to 1961")
autoplot(data_ap)+labs(x="Date", y="Passenger numbers(1000's)", 
                       title = "Air passengers from 1949 to 1961")
boxplot(data_ap~cycle(data_ap), xlab="Date", ylab="Passenger numbers(1000's)",
        main="Monthly Air Passengers from 1949 to 1961")

autoplot(log(data_ap))+labs(x="Date", y="Passenger numbers(1000's)", 
                       title = "Air passengers from 1949 to 1961")
```

**Note** The amplitude of the original data is becoming larger as time goes, so we need to take `log` if want to use `additive` in decomposation.


## Time Series Decomposition
```{r}
decompose_ap <- decompose(data_ap, "multiplicative") 
#decompose_ap'class is decompose.ts
autoplot(decompose_ap)
decompose_ap1 <- decompose(log(data_ap), "additive")
autoplot(decompose_ap1)
```

## Test Stationarity of The Time Series

**1. ADF(Augmented Dickey-Fuller) Test**

```{r}
adf.test(data_ap)
```

**2. ACF(Autocorrelation) Test**

```{r}
autoplot(acf(data_ap, plot=F))+labs(title="Correlogram of Air Passengers from 1949 to 1961")
```

 If the autocorrelation crosses the dashed blue line, it means that specific lag is significantly correlated with current series.
 
```{r}
decompose_ap$random # the first and last 6 residuals are missing
autoplot(acf(decompose_ap$random[7:138],plot=FALSE))+ labs(title="Correlogram of Air Passengers Random Component from 1949 to 1961") 
```


## Fit a Time Series Model

**1. Linear Model**

```{r}
data_ap_dm <- cbind(as.numeric(time(data_ap)),as.numeric(data_ap))
colnames(data_ap_dm) <- c('t','passengers')
ggplot(data_ap_dm, aes(x=t,y=passengers)) + geom_line() + geom_smooth(method = "lm") + labs(x ="Date", y = "Passenger numbers (1000's)", title="Air Passengers from 1949 to 1961") 
```

**2. Arima Model**

`ggtsdiag` is from package `ggfortify` and used to perform diagnostics of residuals.

```{r}
arima_ap <- auto.arima(data_ap)
arima_ap
arima_ap1 <- auto.arima(log(data_ap))
arima_ap1
ggtsdiag(arima_ap)
ggtsdiag(arima_ap1)
```

## Calculate Forecast

```{r}
forecast_ap <- forecast(arima_ap, level=c(95), h=36)
autoplot(forecast_ap)
forecast_ap1 <- forecast(arima_ap1, level=c(95), h=36)
autoplot(forecast_ap1)
as.numeric(forecast_ap1$x)
exp(as.numeric(forecast_ap1$x)) # x is original series
exp(as.numeric(forecast_ap1$mean)) # mean/lower/upper is the predicted series
```


## seasonal adjustment

```{r}
summary(decompose_ap)
autoplot(data_ap/decompose_ap$seasonal)+labs(x="date", y="deseasoned series", title = "original series after deseasoning") 
```

since we choose `multiplicative` in `decompose`, here we need devide by seasonal.

But for logarithm of data, we can use `-`.
```{r}
summary(decompose_ap1)
autoplot(log(data_ap)-decompose_ap1$seasonal)+labs(x="date", y="log of deseasoned series", title = "log of original series after deseasoning") 
```
 
## 

