---
title: "Kalman filter(Partial Pooling)"
author: "LinyiGuo"
date: "2019/8/11"
output:
  pdf_document: default
  html_document: default
---
```{r}
rm(list = ls())
set.seed(9483)
```
Data is from [NIKE](https://www.macrotrends.net/stocks/charts/NKE/nike/revenue) and [ADIDAS](https://www.macrotrends.net/stocks/charts/ADDYY/adidas-ag/revenue).

# Data Understanding & Arima Analysis

```{r}
library(ggfortify)
library(forecast)
library(KFAS)
library(tseries)
library(gridExtra)

# Load Data
data_nike <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\nike.csv',header = TRUE)
data_adi <- read.csv('C:\\Users\\GuoLY\\Desktop\\markdown\\adidas.csv',header = TRUE)

data_nike <- ts(rev(data_nike[3:58,2]), frequency=4, start=c(2005,1))
data_adi <- ts(rev(data_adi[2:57,2]), frequency=4, start=c(2005,1))

# Visualization
autoplot(data_nike) + labs(x='date', y='sales(millionsOfUSD)',
                           title = 'Nike quarterly sales from 2005 to 2018')
autoplot(data_adi) + labs(x='date', y='sales(millionsOfUSD)',
                         title = 'Adidas quarterly sales from 2005 to 2018')

# Time Series Decomposition
decompose_nike <- decompose(data_nike, "additive")
autoplot(decompose_nike)

decompose_adi <- decompose(data_adi, "additive")
autoplot(decompose_adi)

# Fit Arima model
arima_nike <- auto.arima(data_nike)
arima_adi <- auto.arima(data_adi)

# Model diagnostic
ggtsdiag(arima_nike)
ggtsdiag(arima_adi)

# correlation between trends
# two trends have strong positive correlation 0.9211
cor(decompose_nike$trend[3:54], decompose_adi$trend[3:54]) 

# correlation between seasonal
# positive correlation 0.7563
cor(decompose_nike$seasonal,decompose_adi$seasonal)
```

The residuals:

* seem like white noise around 0

* have no significant correlations

* p values are well above 0.05

*Comment:* Ljung-Box is used to test the independence, and the hypothesis being tested is that the residuals from the ARIMA model have no autocorrelation.


# Kalman Filter

## Analyse separately

### NIKE

```{r}

mod_nike <- SSModel(data_nike ~ SSMtrend(1,Q=list(1)) + 
                      SSMseasonal(4,sea.type = "trigonometric"), H = 1)
out_nike <- KFS(mod_nike, filtering = "state")
out_nikeSmooth <- KFS(mod_nike, smoothing = "state")

grid.arrange(
autoplot(ts(rowSums(coef(out_nike, states = "seasonal")[,c(1,3)]),
            frequency = 4, start = c(2005,1))) + 
  labs(title = 'Seasonal component from Kalman Filter', y = 'seaonal'),

autoplot(coef(out_nike, states = "trend")) + 
  labs(title = 'Trend component from Kalman Filter', y = 'trend'),

autoplot(data_nike-rowSums(coef(out_nike)[,c(1,2,4)])) +
  labs(title = 'Residuals from Kalman Filter', y='residuals'),
nrow=3)



# Change variation of Trend from 1 to 0.1
mod_nike1 <- SSModel(data_nike ~ SSMtrend(1,Q=list(.1)) + 
                      SSMseasonal(4,sea.type = "trigonometric"), H = 1)
out_nike1 <- KFS(mod_nike1, filtering = "state")

grid.arrange(
autoplot(ts(rowSums(coef(out_nike1, states = "seasonal")[,c(1,3)]),
            frequency = 4, start = c(2005,1))) + 
  labs(title = 'Seasonal component from Kalman Filter(var_T=0.1)', y = 'seaonal'),

autoplot(coef(out_nike1, states = "trend")) + 
  labs(title = 'Trend component from Kalman Filter(var_T=0.1)', y = 'trend'),

autoplot(data_nike-rowSums(coef(out_nike1)[,c(1,2,4)])) +
  labs(title = 'Residuals from Kalman Filter(var_T=0.1)', y='residuals'),
nrow=3)



# Change the variance of noise in obs eq'n from 1 to 0.1
mod_nike2 <- SSModel(data_nike ~ SSMtrend(1,Q=list(1)) + 
                      SSMseasonal(4,sea.type = "trigonometric"), H = .1)
out_nike2 <- KFS(mod_nike2, filtering = "state")

grid.arrange(
autoplot(ts(rowSums(coef(out_nike2, states = "seasonal")[,c(1,3)]),
            frequency = 4, start = c(2005,1))) + 
  labs(title = 'Seasonal component from Kalman Filter(H=0.1)', y = 'seaonal'),

autoplot(coef(out_nike2, states = "trend")) + 
  labs(title = 'Trend component from Kalman Filter(H=0.1)', y = 'trend'),

autoplot(data_nike-rowSums(coef(out_nike2)[,c(1,2,4)])) +
  labs(title = 'Residuals from Kalman Filter(H=0.1)', y='residuals'),
nrow=3)
```

We can tell from the results above:

* we don't name Q in SSMseasonal -> seasonal is invariant

* decrease variance of noise in obs equation(keep other params' values, the same below) -> Trend will be spickier

* decrease variance of noise of trend -> Trend will be smoother

**Comparison**

```{r}
# Seasonal 
plot(decompose_nike$seasonal,ylab='',ylim=c(-350,450))
par(new=TRUE)
plot(ts(rowSums(coef(out_nike, states = "seasonal")[,c(1,3)]),
            frequency = 4, start = c(2005,1)),
     lty=2, col="green", ylab='',ylim=c(-350,450))
par(new=TRUE)
plot(ts(rowSums(coef(out_nike1, states = "seasonal")[,c(1,3)]),
            frequency = 4, start = c(2005,1)),
     lty=3, col="red", ylab='',ylim=c(-350,450))
par(new=TRUE)
plot(ts(rowSums(coef(out_nike2, states = "seasonal")[,c(1,3)]),
            frequency = 4, start = c(2005,1)),
     lty=4, col="blue", ylab='',ylim=c(-350,450))
title(main = 'Comparison of seasonal', ylab = 'seasonal')
legend('topleft', c('Arima','Q_T=1','Q_T=0.1','H=0.1'),lty=c(1,2,3,4),col=c('black','green','red','blue'),cex=0.5)

# Trend
plot(decompose_nike$trend, ylab='', ylim=c(3000,10000))
par(new=TRUE)
plot(coef(out_nike, states = 'trend'),
          lty=2, col="green", ylab='',ylim=c(3000,10000))
par(new=TRUE)
plot(coef(out_nike1, states = 'trend'),
          lty=3, col="red", ylab='',ylim=c(3000,10000))
par(new=TRUE)
plot(coef(out_nike2, states = 'trend'),
          lty=4, col="blue", ylab='',ylim=c(3000,10000))
title(main = 'Comparison of trend', ylab = 'trend')
legend('topleft', c('Arima','Q_T=1','Q_T=0.1','H=0.1'),lty=c(1,2,3,4),col=c('black','green','red','blue'),cex=0.8)


# error among seasonal and trend

# define mse function
mse <- function(y,y_hat) mean((y-y_hat)^2)

# compute corresponding mse of seasonal
mse_nike_seasonal <- data.frame(
mse(y_hat = rowSums(coef(out_nike, states = "seasonal")[,c(1,3)]),
    y = decompose_nike$seasonal),
mse(y_hat = rowSums(coef(out_nike1, states = "seasonal")[,c(1,3)]),
    y = decompose_nike$seasonal),
mse(y_hat = rowSums(coef(out_nike2, states = "seasonal")[,c(1,3)]),
    y = decompose_nike$seasonal))
colnames(mse_nike_seasonal) <- c('Q=H=1','Q=.1','H=.1')
print(mse_nike_seasonal)

# mse of trend
mse_nike_trend <- data.frame(
  mse(y_hat = coef(out_nike, states = "trend")[3:54], y = decompose_nike$trend[3:54]),
  mse(y_hat = coef(out_nike1, states = "trend")[3:54], y = decompose_nike$trend[3:54]),
  mse(y_hat = coef(out_nike2, states = "trend")[3:54], y = decompose_nike$trend[3:54])
)
colnames(mse_nike_trend) <- c('Q=H=1','Q=.1','H=.1')
print(mse_nike_trend)

```

### ADIDAS(SKIP)




## Partial Pooling

```{r}
mod_combination <- SSModel(cbind(data_nike,data_adi) ~ 
                             SSMtrend(1,Q=list(diag(2))) + 
                             SSMseasonal(4, sea.type = "trigonometric", index=1)+
                             SSMseasonal(4, sea.type = "trigonometric", index=2),
                           H = diag(2))

out_combination <- KFS(mod_combination, filtering = "state", smoothing = "state")
print(out_combination)

# trend
ts.plot(coef(out_combination,states = "trend"), 
        col=c(1,1),lwd=c(1,2), ylim=c(1500,10000), ylab='')
par(new=TRUE)
ts.plot(decompose_nike$trend, col=2, lwd=1, lty=2, ylim=c(1500,10000), ylab='')
par(new=TRUE)
ts.plot(decompose_adi$trend, col=2, lwd=2, lty=2, ylim=c(1500,10000), ylab='')
title(main="Trend component comparison")
legend('topleft', c('Kalman_Nike', 'Kalman_Adi', 'Arima_Nike', 'Arima_Adi'), col=c(1,1,2,2), lwd=c(1,2,1,2))


# seasonal
## Nike
ts.plot(ts(rowSums(coef(out_combination,states="seasonal")[,c(1,3)]),
           frequency = 4, start = c(2005,1)), 
        ylim=c(-350,450), ylab='')
par(new=TRUE)
ts.plot(decompose_nike$seasonal, col=2, lty=2, ylim=c(-350,450), ylab='')
legend("topleft", c('Kalman', 'Arima'), col=c(1,2), lty=c(1,2),cex=0.8)
title(main="Seasonal Components' Comparison of Nike")

## Adi
ts.plot(ts(rowSums(coef(out_combination,states="seasonal")[,c(4,6)]),
           frequency = 4, start = c(2005,1)), 
        ylim=c(-450,600), ylab='')
par(new=TRUE)
ts.plot(decompose_adi$seasonal, col=2, lty=2, ylim=c(-450,600), ylab='')
legend("topleft", c('Kalman', 'Arima'), col=c(1,2), lty=c(1,2),cex=0.8)
title(main="Seasonal Components' Comparison of Adi")

# seasonal adjustment
## Nike
ts.plot(data_nike, ylab='',ylim=c(3000,11000))
par(new=TRUE)
ts.plot(data_nike-decompose_nike$seasonal,
        col=2, lty =2, ylab='',ylim=c(3000,11000))
par(new=TRUE)
ts.plot(data_nike-ts(rowSums(coef(out_combination,states = "seasonal")[,c(1,3)]),
                     frequency = 4, start = c(2005,1)),
        col=3, lty=3, ylab='',ylim=c(3000,11000))
title(main = "Seasonal Adjustment Results(NIKE)")
legend("topleft", c("original","Arima","Kalman"), lty=c(1,2,3), col=c(1,2,3))

## Adi
ts.plot(data_adi, ylab='',ylim=c(1500,7000))
par(new=TRUE)
ts.plot(data_adi-decompose_adi$seasonal,
        col=2, lty =2, ylab='',ylim=c(1500,7000))
par(new=TRUE)
ts.plot(data_adi-ts(rowSums(coef(out_combination,states = "seasonal")[,c(4,6)]),
                     frequency = 4, start = c(2005,1)),
        col=3, lty=3, ylab='',ylim=c(1500,7000))
title(main = "Seasonal Adjustment Results(Adi)")
legend("topleft", c("original","Arima","Kalman"), lty=c(1,2,3), col=c(1,2,3))

```

# Different Situations

## Change the correlation of trend into 1 (correlated)

```{r}
mod_combination1 <- SSModel(cbind(data_nike,data_adi)~
                              SSMtrend(1,Q=list(matrix(c(1,1,1,1),ncol=2)))+ 
                              SSMseasonal(4,sea.type = "trigonometric", index = 1)+
                              SSMseasonal(4,sea.type = "trigonometric", index = 2),
                            H = diag(2))

out_combination1 <- KFS(mod_combination1, filtering = "state", smoothing = "state")
print(out_combination1)

# Trend
## upper line is Nike, the other one is Adidas
ts.plot(coef(out_combination1, states="trend"), ylim=c(1500,10000))
par(new=TRUE)
ts.plot(coef(out_combination, states = "trend"), col=2, lty=2, ylim=c(1500,10000))
legend('topleft', c('Cor=1','Cor=0'),lty=c(1,2),col=c(1,2))
title(main='Comparison of Trend Between Correlated and Uncorrelated')

# Seasonal
## Nike
plot(ts(rowSums(coef(out_combination1,states = "seasonal")[,c(1,3)]),
        frequency = 4, start=c(2005,1)), 
     ylim=c(-350,450), ylab='')
par(new=TRUE)
plot(decompose_nike$seasonal, lty=2, col=2, ylim=c(-350,450),ylab='')
title(main='Comparison of seasonal(NIKE)')
legend('topleft',c('Cor=1','Cor=0'),lty=c(1,2),col=c(1,2),cex=0.8)

plot(ts(rowSums(coef(out_combination1,states = "seasonal")[,c(1,3)]),
        frequency = 4, start=c(2005,1)) - decompose_nike$seasonal,ylab='',
      main='difference of seasonal components Under Cor=1 and Cor=0(NIKE)')

## Adi
plot(ts(rowSums(coef(out_combination1,states = "seasonal")[,c(4,6)]),
        frequency = 4, start=c(2005,1)), 
     ylim=c(-450,600), ylab='')
par(new=TRUE)
plot(decompose_adi$seasonal, lty=2, col=2, ylim=c(-450,600),ylab='')
title(main='Comparison of seasonal(Adi)')
legend('topleft',c('Cor=1','Cor=0'),lty=c(1,2),col=c(1,2),cex=0.8)

plot(ts(rowSums(coef(out_combination1,states = "seasonal")[,c(4,6)]),
        frequency = 4, start=c(2005,1)) - decompose_adi$seasonal,ylab='',
      main='difference of seasonal components Under Cor=1 and Cor=0(Adi)')

```


## Change the covariance matrix(variance) of seasonal

**Note:** The models we used before are all based on invariant seasonal component, now we move on to Gaussian seasonal.

### Working on single case first(NIKE)

**Change the variance of seasonal from 0 to 0.01 (Q=H=1)**

```{r}

mod_nike_seasonal0.01 <- SSModel(data_nike ~ SSMtrend(degree=1,Q=list(1))+
                               SSMseasonal(4,Q=0.01,sea.type = "trigonometric"), 
                      H = 1)
out_nike_seasonal0.01 <- KFS(mod_nike_seasonal0.01,
                             filtering="state",smoothing='state')

# Seasonal
plot(ts(rowSums(coef(out_nike_seasonal0.01,states = "seasonal")[,c(1,3)]),
        frequency = 4, start = c(2005,1)),
     ylab='',ylim=c(-350,450))
par(new=TRUE)
plot(decompose_nike$seasonal,ylim=c(-350,450),col=2,lty=2)
legend("topleft", c('Kalman(Q_sea=0.01)','Arima'), col=c(1,2), lty=c(1,2))
title(main='Seasonal(NIKE)')


# Seasonal Adjustment
plot(data_nike-ts(rowSums(coef(out_nike_seasonal0.01,
                               states = "seasonal")[,c(1,3)]),
        frequency = 4, start = c(2005,1)),
     ylab='', ylim=c(3000,10000))
par(new=TRUE)
plot(data_nike-decompose_nike$seasonal,col='red',lty=2,ylab='',ylim=c(3000,10000))
legend("topleft", c('Kalman(Q_sea=0.01)','Arima'),lty=c(1,2),col=c(1,2))
title(main='Seasonal Adjustment(NIKE)')

# trend
plot(coef(out_nike_seasonal0.01,states = "trend"), ylab='',ylim=c(3000,10000))
par(new=TRUE)
plot(decompose_nike$trend,ylab='',col=2,lty=2,ylim=c(3000,10000))
legend("topleft", c('Kalman(Q_sea=0.01)','Arima'),lty=c(1,2),col=c(1,2))
title(main='Trend (NIKE)')

# residuals
plot(data_nike-rowSums(coef(out_nike_seasonal0.01)),ylim=c(-500,500),ylab='')
par(new=TRUE)
plot(decompose_nike$random, ylim=c(-500,500), ylab='',lty=2,col=2)
legend("topleft", c('kalman(Q_sea=0.01)','Arima'),lty=c(1,2),col=c(1,2))
title(main="Residuals(NIKE)")

```

**variance of seasonal is equal to 1**

```{r}

mod_nike_seasonal1 <- SSModel(data_nike ~ SSMtrend(degree=1,Q=list(1))+
                               SSMseasonal(4,Q=1,sea.type = "trigonometric"), 
                      H = 1)
out_nike_seasonal1 <- KFS(mod_nike_seasonal1,
                             filtering="state",smoothing='state')

# Seasonal
plot(ts(rowSums(coef(out_nike_seasonal1,states = "seasonal")[,c(1,3)]),
        frequency = 4, start = c(2005,1)),
     ylab='',ylim=c(-450,600))
par(new=TRUE)
plot(decompose_nike$seasonal,ylim=c(-450,600),col=2,lty=2)
legend("topleft", c('Kalman(Q_sea=1)','Arima'), col=c(1,2), lty=c(1,2))
title(main='Seasonal(NIKE)')


# Seasonal Adjustment
# which is StatCan's result
plot(data_nike-ts(rowSums(coef(out_nike_seasonal1,
                               states = "seasonal")[,c(1,3)]),
        frequency = 4, start = c(2005,1)),
     ylab='', ylim=c(3000,10000))
par(new=TRUE)
plot(data_nike-decompose_nike$seasonal,col='red',lty=2,ylab='',ylim=c(3000,10000))
legend("topleft", c('Kalman(Q_sea=1)','Arima'),lty=c(1,2),col=c(1,2))
title(main='Seasonal Adjustment(NIKE)')

# trend
plot(coef(out_nike_seasonal1,states = "trend"), ylab='',ylim=c(3000,10000))
par(new=TRUE)
plot(decompose_nike$trend,ylab='',col=2,lty=2,ylim=c(3000,10000))
legend("topleft", c('Kalman(Q_sea=1)','Arima'),lty=c(1,2),col=c(1,2))
title(main='Trend (NIKE)')

# residuals
plot(data_nike-rowSums(coef(out_nike_seasonal1)),ylim=c(-400,500),ylab='')
par(new=TRUE)
plot(decompose_nike$random, ylim=c(-400,500), ylab='',lty=2,col=2)
legend("topleft", c('kalman(Q_sea=1)','Arima'),lty=c(1,2),col=c(1,2))
title(main="Residuals(NIKE)")
```
















