---
  title: "SimulationToFindThePrior"
author: "Linyi Guo"
date: "2020/1/7"
output: 
  html_document: 
    toc: true
---

See my thoughtS in 随笔杂文 2020.1.7.(MY IPAD ONENOTE)

```{r include=FALSE}
set.seed(9483)
rm(list=ls())
```

```{r}
# import package
library(seasonal)
library(forecast)
library(KFAS)
```

# SIMULATION

To avoid the potential mistake, I only simulate the datasets from ARIMA(1,1,1)(0,1,1)


```{r}

model <- Arima(ts(rnorm(120),start=c(1980,01),frequency =12), order=c(1,1,1), seasonal=c(0,1,1),
             fixed=c(phi=runif(1), theta=runif(1), Theta=runif(1)))
data <- simulate(model, nsim=180)
plot(data,type='l')
min(data)


model <- Arima(ts(rnorm(120),start=c(1980,01),frequency =12), order=c(1,1,1), seasonal=c(1,1,1),
             fixed=c(phi=runif(1), theta=runif(1), Phi=runif(1), Theta=runif(1)))
data <- simulate(model, nsim=180)
plot(data,type='l')
min(data)
```

## Define a function to simulate data that we want

```{r}
simulation1 <- function(length){
     
   model <- Arima(ts(rnorm(120),start=c(1980,01),frequency =12), order=c(1,1,1),   
                  seasonal=c(0,1,1), fixed=c(phi=runif(1), theta=runif(1),
                                             Theta=runif(1))
                  )
   
   data <- simulate(model, nsim=length)
   
   # because if we need to take log later, data must be positive
   if(min(data) <= 0) data <- data - min(data) + runif(1)
   else data <- data
   
   return(data)
  
}
```


## function to simulate a datalist with a lot of datasets

```{r}
simlist1 <- function(n,length) {   
  
  Datalist <- list()
  
  for (i in 1:n)  Datalist[[i]] <- simulation1(length)
  
  return(Datalist)  
  
}

```

```{r}
set.seed(9483)

datalist1 <- simlist1(100, 120)
datalist2 <- simlist1(100, 180)
datalist3 <- simlist1(100, 240)
```


Until this step, datalist is basically what I want to use in the following analysis, but we still have one more step: that is to preprocessing.




# Preprocessing

```{r}
# define a function for outliers and log-transformation
preprocess <- function(x11) {
  
  if(transformfunction(x11) == 'log') 
    data <- log(series(x11, 'b1'))
  else
    data <- series(x11, 'b1')

  return(data)
}
```


```{r}
# build model list
x11list1 <- lapply(datalist1, function(x) seas(x, x11=''))
x11list2 <- lapply(datalist2, function(x) seas(x, x11=''))
x11list3 <- lapply(datalist3, function(x) seas(x, x11=''))


```

I can't apply it to a large list(eg. a list with 1000 datasets) because of the following error:

```
Error: X-13 run failed

Errors:
- A model estimation error has occurred during AIC testing within the
  automatic model identification procedure. The error message appears
  below.
- Estimation failed to converge -- maximum iterations reached
```

```{r}
# obtain the preprocessed datalist denoted as Datalist
Datalist1 <- lapply(x11list1, preprocess)
Datalist2 <- lapply(x11list2, preprocess)
Datalist3 <- lapply(x11list3, preprocess)

```

Now we finished this step, then we need to find the 'ideal' values of parameters in these datasets.


# Building Our Prior

## Searching for the 'best' value

**Loss function**

```{r}

Dif1 <- function(x11, ssm, data, sigma){
  
  x11_trend <- series(x11, 'd12')
  x11_seasonal <- series(x11, 'd10')
  x11_irregular <- series(x11, 'd13')
  
  ssm_trend <- coef(ssm, states = 'trend')
  ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
  ssm_irregular <- data[-1] - ssm_trend[-1] - ssm_seasonal[-length(data)]
 
  D <-  sum((x11_irregular[-1]-ssm_irregular)^2)/sigma[1] + 
    sum((x11_trend-ssm_trend)^2)/sigma[2] + 
    sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2)/sigma[3] 
    
  return(D)
}
```

**Exhaustion function**

```{r}

exhaustion1 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
  
   for (i in 1:100) {
     for (j in 1:100) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j*0.1)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i*0.1)
           ssm <- KFS(ssmm)
           
           sigma <- c(i*0.1, j*0.1, 1)
           
           dif <- Dif1(x11, ssm, data, sigma)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```
# system.time(exhaustion1(Datalist1[[1]]))
#   user  system  elapsed
#  189.80 237.43  427.96 

idevallist1 <- lapply(Datalist1, exhaustion1)

```

```
idevalmat1 <- c()
for (i in 1:100){
  
  ideval <- idevallist1[[i]][which.min(idevallist1[[i]]$difference),c(1,2)]
  idevalmat1 <- rbind(idevalmat1, ideval)
  
}

rownames(idevalmat1) <- c(1:100)
head(idevalmat1)

```

```
write.csv(idevalmat1, "C:\\Users\\GuoLY\\Desktop\\markdown\\2020 Winter\\BuildThePrior\\data\\idevalmat1.csv")
# But I changed that to idevalmat1_10.csv later
```

```{r}

idevalmat1_10 <- read.csv("C:\\Users\\GuoLY\\Desktop\\markdown\\2020 Winter\\BuildThePrior\\data\\idevalmat1_10.csv")
```


The result is not very good so I decide to modify the code a little bit and run them again:

```{r}

exhaustion1 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
  
   for (i in 1:100) {
     for (j in 1:100) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j*0.2)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i*0.2)
           ssm <- KFS(ssmm)
           
           sigma <- c(i*0.2, j*0.2, 1)
           
           dif <- Dif1(x11, ssm, data, sigma)
           
           Difference <- c(Difference, dif)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```
system.time(exhaustion1(Datalist1[[1]]))

#用户   系统   流逝 
#189.25 191.60 382.56 
```

```

idevallist1 <- lapply(Datalist1, exhaustion1)
```

```
idevalmat1 <- c()

for (i in 1:100){
  ideval <- idevallist1[[i]][which.min(idevallist1[[i]]$difference),c(1,2)]
  idevalmat1 <- rbind(idevalmat1, ideval)
}


rownames(idevalmat1) <- c(1:100)

```

And then please output the data frame `idevalmat1`.

```
write.csv(idevalmat1, "C:\\Users\\GuoLY\\Desktop\\markdown\\2020 Winter\\BuildThePrior\\data\\idevalmat1_1.csv")
```


```{r}
idevalmat1_1 <- read.csv("C:\\Users\\GuoLY\\Desktop\\markdown\\2020 Winter\\BuildThePrior\\data\\idevalmat1_1.csv")
```


```{r}
hist(idevalmat1_1[,2])

hist(idevalmat1_1[,3])
```


```

set.seed(521)

datalist1 <- simlist1(100, 120)

x11list1 <- lapply(datalist1, function(x) seas(x, x11=''))

Datalist1 <- lapply(x11list1, preprocess)

```

```
idevallist1_2 <- lapply(Datalist1, exhaustion1)
```

```
idevalmat1_2 <- c()

for (i in 1:100){
  ideval <- idevallist1_2[[i]][which.min(idevallist1_2[[i]]$difference),c(1,2)]
  idevalmat1_2 <- rbind(idevalmat1_2, ideval)
}


rownames(idevalmat1_2) <- c(1:100)

```

```
write.csv(idevalmat1_2, "C:\\Users\\GuoLY\\Desktop\\markdown\\2020 Winter\\BuildThePrior\\data\\idevalmat1_2.csv")
```



# Amazing Finding!

I just heard one very useful package named `parallel` and tried it on a toy problem. The result is so amazing! The running time is just one quarter of the previous when I called 4 cores of my computer! Shout out to 老臧！

# Back to game

```

cl <- makeCluster(4)

fun1 <- function(x){
  
  library(seasonal)
  seas(x, x11='') 
  
}

preprocess <- function(x11) {
  
  if(transformfunction(x11) == 'log') 
    data <- log(series(x11, 'b1'))
  else
    data <- series(x11, 'b1')
  
  return(data)
}

#Dif1 <- function(x11, ssm, data, sigma){
#
#  x11_trend <- series(x11, 'd12')
#  x11_seasonal <- series(x11, 'd10')
#  x11_irregular <- series(x11, 'd13')
#  
#  ssm_trend <- coef(ssm, states = 'trend')
#  ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
#  ssm_irregular <- data[-1] - ssm_trend[-1] - ssm_seasonal[-length(data)]
# 
#  D <-  sum((x11_irregular[-1]-ssm_irregular)^2)/sigma[1] + 
#    sum((x11_trend-ssm_trend)^2)/sigma[2] + 
#    sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2)/sigma[3] 
#    
#  return(D)
#}

exhaustion1 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
  
   for (i in 1:100) {
     for (j in 1:100) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j*0.2)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i*0.2)
           ssm <- KFS(ssmm)
           
           sigma <- c(i*0.2, j*0.2, 1)
           
           ### difference ###
           x11_trend <- series(x11, 'd12')
           x11_seasonal <- series(x11, 'd10')
           x11_irregular <- series(x11, 'd13')
           
           ssm_trend <- coef(ssm, states = 'trend')
           ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
           ssm_irregular <- data[-1] - ssm_trend[-1] - ssm_seasonal[-length(data)]
          
           D <-  sum((x11_irregular[-1]-ssm_irregular)^2)/sigma[1] + 
             sum((x11_trend-ssm_trend)^2)/sigma[2] + 
             sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2)/sigma[3] 
           ### end ###
           
           Difference <- c(Difference, D)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}

set.seed(1)

datalist1 <- simlist1(100, 120)

# This is to build the package environment for each core

clusterEvalQ(cl,{
  library(seasonal)
  library(KFAS)
  } 
)


system.time({
  
  x11list1 <- parLapply(cl, datalist1, fun1 )
  
  Datalist1 <- parLapply(cl, x11list1, preprocess )

  idevallist1_3 <- parLapply(cl, Datalist1, exhaustion1)
  
  idevalmat1_3 <- c()
  
  for (i in 1:100){
    ideval <- idevallist1[[i]][which.min(idevallist1[[i]]$difference),c(1,2)]
    idevalmat1_3 <- rbind(idevalmat1_3, ideval)
  }
  
  rownames(idevalmat1_3) <- c(1:100)

  
})

stopCluster(cl)

write.csv(idevalmat1_3, "C:\\Users\\GuoLY\\Desktop\\markdown\\2020 Winter\\BuildThePrior\\data\\idevalmat1_3.csv")
```

# idevalmat1_4

```
rm(list=ls())

library(parallel)
library(forecast)
```
```
# functions

simulation1 <- function(length){
     
   model <- Arima(ts(rnorm(120),start=c(1980,01),frequency =12), order=c(1,1,1),   
                  seasonal=c(0,1,1), fixed=c(phi=runif(1), theta=runif(1),
                                             Theta=runif(1))
                  )
   
   data <- simulate(model, nsim=length)
   
   # because if we need to take log later, data must be positive
   if(min(data) <= 0) data <- data - min(data) + runif(1)
   else data <- data
   
   return(data)
  
}

simlist1 <- function(n,length) {   
  
  Datalist <- list()
  
  for (i in 1:n)  Datalist[[i]] <- simulation1(length)
  
  return(Datalist)  
  
}

fun1 <- function(x){
  
  library(seasonal)
  seas(x, x11='') 
  
}

preprocess <- function(x11) {
  
  if(transformfunction(x11) == 'log') 
    data <- log(series(x11, 'b1'))
  else
    data <- series(x11, 'b1')
  
  return(data)
}



# put previous functions 'exhaustion1' and 'Dif1' together

exhaustion1 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
  
   for (i in 1:100) {
     for (j in 1:100) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j*0.2)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i*0.2)
           ssm <- KFS(ssmm)
           
           sigma <- c(i*0.2, j*0.2, 1)
           
           ### difference ###
           x11_trend <- series(x11, 'd12')
           x11_seasonal <- series(x11, 'd10')
           x11_irregular <- series(x11, 'd13')
           
           ssm_trend <- coef(ssm, states = 'trend')
           ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
           ssm_irregular <- data[-1] - ssm_trend[-1] - ssm_seasonal[-length(data)]
          
           D <-  sum((x11_irregular[-1]-ssm_irregular)^2)/sigma[1] + 
             sum((x11_trend-ssm_trend)^2)/sigma[2] + 
             sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2)/sigma[3] 
           ### end ###
           
           Difference <- c(Difference, D)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```

set.seed(1)
datalist1 <- simlist1(100, 120)

cl <- makeCluster(4)

clusterEvalQ(cl,{
  library(seasonal)
  library(KFAS)
  } 
)

system.time({
  
  x11list1 <- parLapply(cl, datalist1, fun1 )
  
  Datalist1 <- parLapply(cl, x11list1, preprocess )

  idevallist1_4 <- parLapply(cl, Datalist1, exhaustion1)
  
  idevalmat1_4 <- c()
  
  for (i in 1:100){
    ideval <- idevallist1_4[[i]][which.min(idevallist1_4[[i]]$difference),c(1,2)]
    idevalmat1_4 <- rbind(idevalmat1_4, ideval)
  }
  
  rownames(idevalmat1_4) <- c(1:100)

})

stopCluster(cl)

```

# idevalmat3_3

```
# functions

simulation1 <- function(length){
     
   model <- Arima(ts(rnorm(120),start=c(1980,01),frequency =12), order=c(1,1,1),   
                  seasonal=c(0,1,1), fixed=c(phi=runif(1), theta=runif(1),
                                             Theta=runif(1))
                  )
   
   data <- simulate(model, nsim=length)
   
   # because if we need to take log later, data must be positive
   if(min(data) <= 0) data <- data - min(data) + runif(1)
   else data <- data
   
   return(data)
  
}

simlist1 <- function(n,length) {   
  
  Datalist <- list()
  
  for (i in 1:n)  Datalist[[i]] <- simulation1(length)
  
  return(Datalist)  
  
}

fun1 <- function(x){
  
  library(seasonal)
  seas(x, x11='') 
  
}

preprocess <- function(x11) {
  
  if(transformfunction(x11) == 'log') 
    data <- log(series(x11, 'b1'))
  else
    data <- series(x11, 'b1')
  
  return(data)
}



# put previous functions 'exhaustion1' and 'Dif1' together

exhaustion1 <- function(data){
  
  Difference <- c()
  index <- c()
  
  x11 <- seas(data, x11='')
  
   for (i in 1:100) {
     for (j in 1:100) {
         
           ssmm <- SSModel(data ~ SSMtrend(1, Q=list(j*0.2)) + 
                     SSMseasonal(12, sea.type = 'dummy', Q = 1),
                   H = i*0.2)
           ssm <- KFS(ssmm)
           
           sigma <- c(i*0.2, j*0.2, 1)
           
           ### difference ###
           x11_trend <- series(x11, 'd12')
           x11_seasonal <- series(x11, 'd10')
           x11_irregular <- series(x11, 'd13')
           
           ssm_trend <- coef(ssm, states = 'trend')
           ssm_seasonal <- -rowSums(coef(ssm, states='seasonal'))
           ssm_irregular <- data[-1] - ssm_trend[-1] - ssm_seasonal[-length(data)]
          
           D <-  sum((x11_irregular[-1]-ssm_irregular)^2)/sigma[1] + 
             sum((x11_trend-ssm_trend)^2)/sigma[2] + 
             sum((x11_seasonal[-1]-ssm_seasonal[-length(data)])^2)/sigma[3] 
           ### end ###
           
           Difference <- c(Difference, D)
           
           index <- rbind(index, sigma)
      }
   }
  
  df <- data.frame(variance=index, difference = Difference)
  return(df)
}
```

```

set.seed(1325341)
datalist3 <- simlist1(100, 240)

cl <- makeCluster(6)

clusterEvalQ(cl,{
  library(seasonal)
  library(KFAS)
  } 
)

system.time({
  
  x11list3 <- parLapply(cl, datalist3, fun1 )
  
  Datalist3 <- parLapply(cl, x11list3, preprocess )

  idevallist3_3 <- parLapply(cl, Datalist3, exhaustion1)
  
  idevalmat3_3 <- c()
  
  for (i in 1:100){
    ideval <- idevallist3_3[[i]][which.min(idevallist3_3[[i]]$difference),c(1,2)]
    idevalmat3_3 <- rbind(idevalmat3_3, ideval)
  }
  
  rownames(idevalmat3_3) <- c(1:100)

})

stopCluster(cl)

write.csv(idevalmat3_3, "C:\\Users\\GuoLY\\Desktop\\markdown\\2020 Winter\\BuildThePrior\\data\\idevalmat3_3.csv")

```