---
title: "meeting report 5.27"
author: "linyiguo"
date: "2019-5-27"
output:
  word_document: default

---
Dear Aaron, 

I am writing to give you a brief about the stuff that I have done in last three days.

# I Simulation
I tried to simulate data from a fixed seasonal arima model, the code is mainly referred to [this website](https://robjhyndman.com/hyndsight/simulating-from-a-specified-seasonal-arima-model/). Of course I tried some other methods as well, but this one seems to be more reliable. But the questions is, I think the data simulated is kinda of not right.
```{r}
library(forecast)
set.seed(1)
model <- Arima(ts(rnorm(24000),freq=12), order=c(0,1,1), seasonal=c(0,1,1),fixed=c(theta=0.5, Theta=0.5))
foo <- simulate(model,nsim = 240)
plot(foo,type="l")
fit <- Arima(foo, order=c(0,1,1), seasonal=c(0,1,1))
summary(fit)
```

# II Reproduction

The main reference is the [document](https://cran.r-project.org/web/packages/seasonal/vignettes/seas.pdf) about *X-13ARIMA-SEATS(X-13)*. Something to clearify: 
By default,  a call to **seas** also invokes the following automatic procedures of X-13:

* Transformation selection (log / no log);   
* Detection of trading day and Easter effects;   
* Outlier detection;   
* ARIMA model search.   

By default, seas calls the *SEATS* adjustment procedure(which decomposes the ARIMA model).To perform the alternative *X-11* adjustment procedure, we need to add **x11 = " "**.  

But when I tried to use these code on the simulated data, the curves I got from them are too smooth and looks same. I am thinking: maybe the data I simulated before is not appropriate. In addition, something wrong with the SEATS, cause the model from it(*SARIMA(0,1,1)(0,1,0)[12]*) is different from that of x-11, which is close to our true model *SARIMA(0,1,1)(0,1,1)[12]* 
```{r}
library(seasonal)
library(forecast)
set.seed(1)
model <- Arima(ts(rnorm(24000),freq=12), order=c(0,1,1), seasonal=c(0,1,1),fixed=c(theta=0.5, Theta=0.5))
data <- simulate(model,nsim=240)
plot(data)
m_x11 <- seas(data, x11 = "", regression.aictest =  NULL)
plot(m_x11)
m_seats <- seas(data, regression.aictest = NULL)
plot(m_seats)

plot(data,col="green",ylim=c(-10,2000),ylab="")
par(new=T)
plot(final(m_x11),col="red",ylim=c(-10,2000),ylab="")
par(new=T)
plot(final(m_seats),col="blue",ylim=c(-10,2000),ylab="")
legend("topleft",c("Data","X11","SEATS"),col=c("green","red","blue"),lty=c(1,1,1))

```

I tried the same code on the data set **unemp**, whose results seem to be good, at least not very smooth and not totally same:
```{r}
library(seasonal)
library(seasonalview)
eg_seats <- seas(unemp)
eg_x11 <- seas(unemp, x11 = "")
plot(unemp,col="lightblue",ylim=c(5500,15300),ylab="")
par(new=T)
plot(final(eg_seats),col="red",ylim=c(5500,15300),ylab="")
par(new=T)
plot(final(eg_x11),col="blue",ylim=c(5500,15300),ylab="")
```


And I am still working on the state space model.

**Update 5.30**
以下是Aaron的回复，包含了几条不错的建议：

* The discuss about 'spickness' is not enough, to verify *SEAS* can decompose our seasonal/trend component, we can try: i) remove the S/T from our data, which can be achieved only when data is simulated, cause we know the specific model for each component; ii) detrend/deseason our data by *seas*, and compare two series to see whether *seas* works well(yes if both seem similar)

* the seasonal component we used before is kinda of easy or regular, maybe we can try some more complicated case? like kinds of holidays.

* In reality, the noise is always not gaussian, so maybe we can create some our own noise. To be specific, use the residuals of one data set, like **unemp**. The residual is just the true value minus the prediction of some reasonable model. And we can build a 'noise' library, this may be helpful in future.



