---
title: "reproducing2"
author: "linyiguo"
date: "2019.6.20"
output: word_document
---

--------------------------------------**UPDATE 6.20**--------------------------------------     

中间隔了很久， 很多内容在之间的几次meeting report里，接下来要做的事情，主要是根据上一次meeting *6.17*的内容。

### trend$sd=5, season fixed, weight$sd=10*k(k=range(10))
```{r}
Sys.setlocale('LC_ALL','C')
library(seasonal)
library(forecast)
set.seed(1)

# generate data
model <- Arima(ts(rnorm(24000),freq=12), order=c(0,1,1), seasonal=c(0,1,1),fixed=c(theta=0.5, Theta=0.5))
data <- simulate(model,nsim=240)
plot(data,type="l")

# define m_x11
m_x11 <- seas(data, x11 = "", regression.aictest =  NULL)
plot(final(m_x11),type="l",ylim=c(-50,2000),ylab="")
par(new=T)
plot(as.numeric(data-final(m_x11)), type="l",ylim=c(-50,2000),ylab="")

# Initialization
a = 150 # a-12 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (k in 1:10) {
  component <- c() # reset the component after each case
  S <- matrix(0,1000,252) 
  Tr <- matrix(0,1000,252)
  Tr[,12] <- data[1]
  S[,1:12] <- matrix(as.numeric(data-final(m_x11))[1:12],nrow=1000,ncol=12,byrow=T)
  for (i in 13:a) {
    # update particles
    Tr[,i] <- Tr[,i-1] + rnorm(1000,sd=5)
    if(i%%12==0) S[,i]=S[,12]
    else S[,i]=S[,i%%12]
  
    # update weights
    w <- dnorm(data[i-12]-Tr[,i]-S[,i],sd=10*k)
    w <- w/sum(w)
    
    # evaluate state value
    t <- sum(w * Tr[,i])
    s <- sum(w * S[,i])
    
    # add to our component path
    component <- rbind(component, c(t,s))
    
    # resample
    Tr[,i] <- sample(Tr[,i], size =1000, replace = TRUE, prob = w)
    S[,i] <- sample(S[,i], size = 1000, replace = TRUE, prob = w)
  }
  
  # data vs deseasoned
  plot(data[1:(a-12)],type = "l",ylim = c(lower,upper),ylab='')
  par(new=TRUE)
  plot(data[1:(a-12)]-component[,2],type="l",col="red",ylim=c(lower,upper),ylab='')
  par(new=TRUE)
  plot(component[,2],type="l",col="blue",ylim=c(lower,upper),ylab='')
  legend("topleft", c("data","deseasoned"),col=c("black","red"),lty=c(1,1))

}

```
不是很清楚这个warning是什么意思，这个可能就是导致我们这里的seasonal component不固定的原因。(**SOLVED**)

I realized there is no meaning to do such a plot! Our season is fixed so deseasoned curves must be same for these cases. Actually, we should try the curve of *trend+season* since currently the independent variable we focus on is the variance of our weight, which represents the variance of data-(trend+season),that is the distance between data and trend+season:

### trend$sd=5, season fixed, weight$sd=10*k(k=range(10))
```{r}


# Initialization
a = 150 # a-12 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (k in 1:10) {
  component <- c() # reset the component after each case
  S <- matrix(0,1000,252) 
  Tr <- matrix(0,1000,252)
  Tr[,12] <- data[1]
  S[,1:12] <- matrix(as.numeric(data-final(m_x11))[1:12],nrow=1000,ncol=12,byrow=T)
  for (i in 13:a) {
    # update particles
    Tr[,i] <- Tr[,i-1] + rnorm(1000,sd=5)
    if(i%%12==0) S[,i]=S[,12]
    else S[,i]=S[,i%%12]
  
    # update weights
    w <- dnorm(data[i-12]-Tr[,i]-S[,i],sd=10*k)
    w <- w/sum(w)
    
    # evaluate state value
    t <- sum(w * Tr[,i])
    s <- sum(w * S[,i])
    
    # add to our component path
    component <- rbind(component, c(t,s))
    
    # resample
    Tr[,i] <- sample(Tr[,i], size =1000, replace = TRUE, prob = w)
    S[,i] <- sample(S[,i], size = 1000, replace = TRUE, prob = w)
  }
  
  # data vs deseasoned
  plot(data[1:(a-12)],type = "l",ylim = c(lower,upper),ylab='')
  par(new=TRUE)
  plot(component[,1]+component[,2],type="l",col="red",ylim=c(lower,upper),ylab='')
  par(new=TRUE)
  plot(component[,1],type="l",col="blue",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(data[1:(a-12)]-component[,1]-component[,2],type="l",col='green',ylim=c(lower,upper),ylab='')
  legend("topleft", c("data","T+S","T",'detrend&deseason'),col=c("black","red","blue",'green'),lty=c(1,1,1,1))

}


```

**comment: ** The trend+season curve here is not smooth enough, but our trend is very smooth. We should try to let seasonal part be changeable.

--------------------------------------------------
**UPDATE 6.22** 自己似乎忘记了我们的重点是要哪条曲线。不忘初衷啊年轻人。seasonal adjustment！我们最终是想估计季节变化的影响并把这个影响从现实数据中移除。因此还是要先关注seasonal adjustment之后的数据，其它的可以先放一放。 
--------------------------------------------------


### trend$sd=5, season$sd=5, weight$sd=10*k(k=range(10))
```{r}
Sys.setlocale('LC_ALL','C')
library(seasonal)
library(forecast)
set.seed(1)

# generate data
model <- Arima(ts(rnorm(24000),freq=12), order=c(0,1,1), seasonal=c(0,1,1),fixed=c(theta=0.5, Theta=0.5))
data <- simulate(model,nsim=240)

# define m_x11
m_x11 <- seas(data, x11 = "", regression.aictest =  NULL)

# Initialization
a = 150 # a-12 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (k in 0:10) {
  component <- c() # reset the component after each case
  S <- matrix(0,1000,252) 
  Tr <- matrix(0,1000,252)
  Tr[,12] <- data[1]
  S[,1:12] <- matrix(as.numeric(data-final(m_x11))[1:12],nrow=1000,ncol=12,byrow=T)
  for (i in 13:a) {
    # update particles
    Tr[,i] <- Tr[,i-1] + rnorm(1000,sd=5)
    for (j in 1:11) S[,i] <- S[,i]-S[,i-j]
    S[,i] <- S[,i] + rnorm(1000,sd=5)
  
    # update weights
    w <- dnorm(data[i-12]-Tr[,i]-S[,i],sd=5+10*k)
    w <- w/sum(w)
    
    # evaluate state value
    t <- sum(w * Tr[,i])
    s <- sum(w * S[,i])
    
    # add to our component path
    component <- rbind(component, c(t,s))
    
    # resample
    Tr[,i] <- sample(Tr[,i], size =1000, replace = TRUE, prob = w)
    S[,i] <- sample(S[,i], size = 1000, replace = TRUE, prob = w)
  }
  
  # data vs deseasoned
  plot(data[1:(a-12)],type = "l",ylim = c(lower,upper),ylab='')
  par(new=T)
  plot(data[1:(a-12)]-component[,2],type="l",col="orange",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,2],type="l",col="blue",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,1],type="l",col="red",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(data-final(m_x11))[1:(a-12)],type="l",lty=2,ylim=c(lower,upper),ylab='')
  legend("topleft", c("data","deseasoned","trend","season"),col=c("black","orange","red","blue"),lty=c(1,1,1,1))

}

```


### 我们在权重的计算中加入对seasonal的考虑(# of particles = 100000)
**trend$sd = season$sd = 5**
```{r}

# Initialization
a = 180 # a-12 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (k in 0:10) {
  component <- c() # reset the component after each case
  S <- matrix(0,100000,252) 
  Tr <- matrix(0,100000,252)
  Tr[,12] <- data[1]
  S[,1:12] <- matrix(as.numeric(data-final(m_x11))[1:12],nrow=100000,ncol=12,byrow=T)
  for (i in 13:a) {
    # update particles
    Tr[,i] <- Tr[,i-1] + rnorm(100000,sd=5)
    for (j in 1:11) S[,i] <- S[,i]-S[,i-j]
    S[,i] <- S[,i] + rnorm(100000,sd=5)
  
    # update weights
    w <- dnorm(data[i-12]-Tr[,i]-S[,i],sd=5+10*k) * dnorm(rowSums(S[,(i-11):i]),sd=5)
    w <- w/sum(w)
    
    # evaluate state value
    t <- sum(w * Tr[,i])
    s <- sum(w * S[,i])
    
    # add to our component path
    component <- rbind(component, c(t,s))
    
    # resample
    Tr[,i] <- sample(Tr[,i], size =100000, replace = TRUE, prob = w)
    S[,i] <- sample(S[,i], size = 100000, replace = TRUE, prob = w)
  }
  
  # data vs deseasoned
  plot(data[1:(a-12)],type = "l",ylim = c(lower,upper),ylab='')
  par(new=T)
  plot(data[1:(a-12)]-component[,2],type="l",col="orange",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,2],type="l",col="blue",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,1],type="l",col="red",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(data-final(m_x11))[1:(a-12)],type="l",lty=2,ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(final(m_x11))[1:(a-12)],type="l",lty=2,col="purple",ylim=c(lower,upper),ylab='')
  legend("topleft", c("data","deseasoned","trend","season"),col=c("black","orange","red","blue"),lty=c(1,1,1,1))

}

```

### Back to our simulated data, use conventional methods to get season/trend and the irregular.

```{r}
library(tseries)
library(seasonalview)
Sys.setlocale('LC_ALL','C')
unemp_x11=seas(unemp,x11="")
summary(unemp_x11)
unemp
m = seas(unemp)
series(m,"forecast.forecasts")

plot(series(m,"forecast.forecasts"))
series(m,c("forecast.forecasts","s12"))
series(m,"history.saestimates")
series(m,"slidingspans.sfspans")
# out(m)  shows the content of the main X-13 output in the browser
# which gives access to every detail of the estimation
view(m)
view(m_x11)

# use the initial code we learned at the beginning
dif1data <- diff(data,differences = 1)
plot(dif1data)
adf.test(dif1data,alternative = "stationary")
library("TSA")
PGram<-periodogram(dif1data)
PGramAsDataFrame = data.frame(freq=PGram$freq, spec=PGram$spec)
order = PGramAsDataFrame[order(-PGramAsDataFrame$spec),]
top2 = head(order)
top2
1/top2[1,1]
datafreq12 <- ts(data,freq=12)
datadecom<-decompose(datafreq12)
plot(datadecom)
plot(datadecom$seasonal,ylim=c(-60,60),col="red")
par(new=T)
plot(data-final(m_x11),ylim=c(-60,60))

```


## treat 24 seasonal components as a cycle

Recall that our former formula of seasonal component is $S_t = \sum_{j=1}^{11} S_{t-j}+\epsilon_t$; and now we use $S_t = \sum_{j=1}^{23} S_{t-j}+\epsilon_t$, that is, we consider more states before when we decide current state.

** trend$sd = season$sd = 5, # of particles = 100000, weight$sd=5+10*k **
```{r}

Sys.setlocale('LC_ALL','C')
# Initialization
a = 120 # a-23 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (k in 0:10) {
  component <- c() # reset the component after each case
  S <- matrix(0,100000,263) 
  Tr <- matrix(0,100000,263)
  Tr[,23] <- data[1]
  S[,1:23] <- matrix(as.numeric(data-final(m_x11))[1:23],nrow=100000,ncol=23,byrow=T)
  for (i in 24:a) {
    # update particles
    Tr[,i] <- Tr[,i-1] + rnorm(100000,sd=5)
    for (j in 1:23) S[,i] <- S[,i]-S[,i-j]
    S[,i] <- S[,i] + rnorm(100000,sd=5)
  
    # update weights
    w <- dnorm(data[i-23]-Tr[,i]-S[,i],sd=5+10*k)
    w <- w/sum(w)
    
    # evaluate state value
    t <- sum(w * Tr[,i])
    s <- sum(w * S[,i])
    
    # add to our component path
    component <- rbind(component, c(t,s))
    
    # resample
    Tr[,i] <- sample(Tr[,i], size =100000, replace = TRUE, prob = w)
    S[,i] <- sample(S[,i], size = 100000, replace = TRUE, prob = w)
  }
  
  # data vs deseasoned
  plot(data[1:(a-23)],type = "l",ylim = c(lower,upper),ylab='')
  par(new=T)
  plot(data[1:(a-23)]-component[,2],type="l",col="orange",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,2],type="l",col="blue",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,1],type="l",col="red",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(data-final(m_x11))[1:(a-23)],type="l",lty=2,ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(final(m_x11))[1:(a-23)],type="l",lty=2,col="purple",ylim=c(lower,upper),ylab='')
  legend("topleft", c("data","deseasoned","trend","season"),col=c("black","orange","red","blue"),lty=c(1,1,1,1))

}
```
## treat 24 seasonal components as a cycle and consider season influence in weight expression

** trend$sd = season$sd = 5, # of particles = 100000**

```{r}
Sys.setlocale('LC_ALL','C')
library(seasonal)
# Initialization
a = 110 # a-23 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (k in 0:10) {
  component <- c() # reset the component after each case
  S <- matrix(0,100000,263) 
  Tr <- matrix(0,100000,263)
  Tr[,23] <- data[1]
  S[,1:23] <- matrix(as.numeric(data-final(m_x11))[1:23],nrow=100000,ncol=23,byrow=T)
  for (i in 24:a) {
    # update particles
    Tr[,i] <- Tr[,i-1] + rnorm(100000,sd=5)
    for (j in 1:23) S[,i] <- S[,i]-S[,i-j]
    S[,i] <- S[,i] + rnorm(100000,sd=5)
  
    # update weights
    w <- dnorm(data[i-23]-Tr[,i]-S[,i],sd=5+10*k) * dnorm(rowSums(S[,(i-23):i]),sd=5)
    w <- w/sum(w)
    
    # evaluate state value
    t <- sum(w * Tr[,i])
    s <- sum(w * S[,i])
    
    # add to our component path
    component <- rbind(component, c(t,s))
    
    # resample
    Tr[,i] <- sample(Tr[,i], size =100000, replace = TRUE, prob = w)
    S[,i] <- sample(S[,i], size = 100000, replace = TRUE, prob = w)
  }
  
  # data vs deseasoned
  plot(data[1:(a-23)],type = "l",ylim = c(lower,upper),ylab='')
  par(new=T)
  plot(data[1:(a-23)]-component[,2],type="l",col="orange",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,2],type="l",col="blue",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,1],type="l",col="red",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(data-final(m_x11))[1:(a-23)],type="l",lty=2,ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(final(m_x11))[1:(a-23)],type="l",lty=2,col="purple",ylim=c(lower,upper),ylab='')
  legend("topleft", c("data","deseasoned","trend","season"),col=c("black","orange","red","blue"),lty=c(1,1,1,1))

}
```

## Seasonal cycle = 24 

**trend$sd=5, season$sd = 1, weight$sd=5+10*k **
```{r}
library(seasonal)
# Initialization
a = 160 # a-23 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (k in 0:10) {
  component <- c() # reset the component after each case
  S <- matrix(0,100000,263) 
  Tr <- matrix(0,100000,263)
  Tr[,23] <- data[1]
  S[,1:23] <- matrix(as.numeric(data-final(m_x11))[1:23],nrow=100000,ncol=23,byrow=T)
  for (i in 24:a) {
    # update particles
    Tr[,i] <- Tr[,i-1] + rnorm(100000,sd=5)
    for (j in 1:23) S[,i] <- S[,i]-S[,i-j]
    S[,i] <- S[,i] + rnorm(100000,sd=1)
  
    # update weights
    w <- dnorm(data[i-23]-Tr[,i]-S[,i],sd=5+10*k)
    w <- w/sum(w)
    
    # evaluate state value
    t <- sum(w * Tr[,i])
    s <- sum(w * S[,i])
    
    # add to our component path
    component <- rbind(component, c(t,s))
    
    # resample
    Tr[,i] <- sample(Tr[,i], size =100000, replace = TRUE, prob = w)
    S[,i] <- sample(S[,i], size = 100000, replace = TRUE, prob = w)
  }
  
  # data vs deseasoned
  plot(data[1:(a-23)],type = "l",ylim = c(lower,upper),ylab='')
  par(new=T)
  plot(data[1:(a-23)]-component[,2],type="l",col="orange",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,2],type="l",col="blue",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,1],type="l",col="red",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(data-final(m_x11))[1:(a-23)],type="l",lty=2,ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(final(m_x11))[1:(a-23)],type="l",lty=2,col="purple",ylim=c(lower,upper),ylab='')
  legend("topleft", c("data","deseasoned","trend","season"),col=c("black","orange","red","blue"),lty=c(1,1,1,1))

}
```



## Seasonal cycle = 24 and consider the seasonal influence on weights

**trend$sd=5, season$sd = 1, weight$sd=5+10*k **
```{r}
library(seasonal)
Sys.setlocale('LC_ALL','C')
# Initialization
a = 150 # a-23 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (k in 0:10) {
  component <- c() # reset the component after each case
  S <- matrix(0,100000,263) 
  Tr <- matrix(0,100000,263)
  Tr[,23] <- data[1]
  S[,1:23] <- matrix(as.numeric(data-final(m_x11))[1:23],nrow=100000,ncol=23,byrow=T)
  for (i in 24:a) {
    # update particles
    Tr[,i] <- Tr[,i-1] + rnorm(100000,sd=5)
    for (j in 1:23) S[,i] <- S[,i]-S[,i-j]
    S[,i] <- S[,i] + rnorm(100000,sd=1)
  
    # update weights
    w <- dnorm(data[i-23]-Tr[,i]-S[,i],sd=5+10*k) * dnorm(rowSums(S[,(i-23):i]),sd=1)
    w <- w/sum(w)
    
    # evaluate state value
    t <- sum(w * Tr[,i])
    s <- sum(w * S[,i])
    
    # add to our component path
    component <- rbind(component, c(t,s))
    
    # resample
    Tr[,i] <- sample(Tr[,i], size =100000, replace = TRUE, prob = w)
    S[,i] <- sample(S[,i], size = 100000, replace = TRUE, prob = w)
  }
  
  # data vs deseasoned
  plot(data[1:(a-23)],type = "l",ylim = c(lower,upper),ylab='')
  par(new=T)
  plot(data[1:(a-23)]-component[,2],type="l",col="orange",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,2],type="l",col="blue",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(component[,1],type="l",col="red",ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(data-final(m_x11))[1:(a-23)],type="l",lty=2,ylim=c(lower,upper),ylab='')
  par(new=T)
  plot(as.numeric(final(m_x11))[1:(a-23)],type="l",lty=2,col="purple",ylim=c(lower,upper),ylab='')
  legend("topleft", c("data","deseasoned","trend","season"),col=c("black","orange","red","blue"),lty=c(1,1,1,1))

}
```