---
title: "simulation exercise"
author: "linyiguo"
date: "2019.5.24"
output:
  word_document: default
  html_document: default
---
我在想借助**R Markdown**这个工具来记录我平时学习的心路历程。在做研究的过程中，势必会产生很多感想，不论是感性上还是理性上，我都愿意把这些ideas记录下来，这样也许更有助于我理清自己的思路，不至于脑子里总是一盆浆糊。在以后，我可能更想用英语来写，但目前个人能力有限，就慢慢来吧。

After meeting with Aaron，I think I still need to spend more time on research, although I have other plans on working out and ielts. But research should be my first goal, given my current situation. Well, let's say what we were talking about in this morning:
I am trying to simulate some data from different models and these data will be used in my later research(since I do not have real data now).

# Auto-regressive model
Let's say we want to simulate data corresponding to model *AR(1)*:$$Z_t = 0.9*Z_{t-1}+\epsilon_t$$
where $\epsilon\sim N(0,1)$, $Z_0=1$

```{r}
set.seed(1)
epsilon <- rnorm(1000)
Z <- rep(0,1001)
Z[1] <- 1
for(i in 2:1001)  Z[i] <- 0.9*Z[i-1] + epsilon[i-1]
plot(Z[2:1001], type="l")
```
```{r}
set.seed(1)
epsilon <- rnorm(1000)
Z <- rep(0,1001)
Z[1] <- 100
for(i in 2:1001)  Z[i] <- 0.9*Z[i-1] + epsilon[i-1]
plot(Z[2:1001], type="l")
```

```{r}
set.seed(1)
Z <- rep(0,1001)
Z[1] <- 1
for(i in 2:1001)  Z[i] <- 0.01*Z[i-1] + epsilon[i-1]
plot(Z[2:1001], type="l")
```
```{r}
set.seed(1)
Z <- rep(0,5001)
e <- rnorm(5000)
Z[1] <- 1
for(i in 2:5001)  Z[i] <- 0.9*Z[i-1] + e[i-1]
plot(Z[2:5001], type="l")
```

**comment:** 从上面的图像可以看出，\phi的值越大，振幅就越大；同时，初始值$Z_0$不会对最终的结果产生影响。我在想一个问题，为什么$\phi_1$的值和ts的amplitude有关？Intuitively，$Z_{t-1}$ is also a variable, which has a normal dist'n as well(since we assume noises are normal), so $Z_t$ is the sum of several normal dist'n, but is mainly determined by the first ones(since $\phi<0$, efficients converge to 0).

# Moving-average model
Here, we simulate data to fit the model *MA(1)*:$$Z_t = \epsilon_t + 0.9*\epsilon_{t-1}$$
where $\epsilon \sim N(0,1)$, $Z_0=1$

```{r}
set.seed(1)
Z <- rep(0,1001)
epsilon <- rnorm(1001)
Z[1] <- 0
for(i in 2:1001) Z[i] = epsilon[i]+0.9*epsilon[i-1]
plot(Z[2:1001],type="l")
```

# Auto-regressive moving-average model
Let's say we want to simulate data according to an *ARMA(1,1)* model:$$Z_t-0.9*Z_{t-1}=\epsilon_t+0.9*\epsilon_{t-1}$$
where $\epsilon \sim N(0,1)$, $Z_0=1$
```{r}
set.seed(1)
Z <- rep(0,2001)
e <- rnorm(2000)
for(i in 2:2001) Z[i] = 0.9*Z[i-1]+e[i]+0.9*e[i-1]
plot(Z,type="l")
```

**[update 2019.5.26]** 其实我发现simulate数据并不是我想的这么麻烦，在R中有一些code可以帮助我们很轻易地模拟得到想要的数据，比如：
```{r}
a <- arima.sim(model=list(ar=c(0.9),ma=c(0.9)), n=1000,innov = rnorm(1000))
plot(a)
```

# Auto-regressive integrated moving-average model
Suppose we want to simulate data to fit *ARIMA(1,1,1)* model:$$(1-0.9B)(1-B)Z_t=\epsilon_t+0.9*\epsilon_{t-1}$$
where $\epsilon\sim N(0,1)$, $Z_0=Z_1=1$
```{r}
set.seed(1)
Z <- rep(0,1001)
e <- rnorm(1000)
Z[1] <- 1
Z[2] <- 1
for (i in 3:1001) Z[i] <- 1.9*Z[i-1]-0.9*Z[i-2]+e[i-1]+0.9*e[i-2]
plot(Z[3:1001],type="l")
Y <- Z[2:1001] - Z[1:1000]
plot(Y, type="l")
```

**comment**: 

# Seasonal Autoregressive Moving-average Model
Now, we consider *SARIMA* model， which can be viewed as an expanded model of *ARIMA*. Let's say our model is $ARIMA(1,1,1)*(1,1,1)_4$, that is,$$(1-0.9B)(1-0.9B^4)(1-B)(1-B^4)Z_t=(1-0.9B)(1-0.9B^4)\epsilon_t$$
After expansion, we have $$(1-1.9B+0.9B^2-1.9B^4+3.61B^5-1.71B^6+0.9B^8-1.71B^9+0.81B^10)*Z_t=(1-0.9B-0.9B^4+0.81B^5)*\epsilon_t$$
where$\epsilon \sim N(0,1)$, we take $Z_1,Z_2,...,Z_{10}$ randomly from normal distribution N(0,1).
```{r}
set.seed(1)
Z <- rep(0,1001)
e <- rnorm(1001)
#Z[1:10] <- rnorm(10)
Z[1:10] <- c(1,2,3,4,1,2,3,4,1,2)
for (i in 11:1001) {
  Z[i] <- e[i]-0.9*e[i-1]-0.9*e[i-4]+0.81*e[i-5]+1.9*Z[i-1]-0.9*Z[i-2]+1.9*Z[i-4]-3.61*Z[i-5]+1.71*Z[i-6]-0.9*Z[i-8]+1.71*Z[i-9]-0.81*Z[i-10]
}
plot(Z[11:1001],type="l")
Y <- Z[6:1001]-Z[5:1000]-Z[2:997]+Z[1:996]
plot(Y,type="l")
plot(Z[1:10],type="l")
```

Well, this is a little weird, cause the curve I expect should be with obvious seasonal fluctuations. Let's try another *SARIMA* model: $$(1-B)(1-B^4)Z_t=(1-0.4B)(1-0.6B^4)\epsilon_t$$
which is equal to $$Z_t=Z_{t-1}+Z_{t-4}-Z_{t-5}+\epsilon_t-0.4*\epsilon_{t-1}-0.6*\epsilon_{t-4}+0.24*\epsilon_{t-5}$$
```{r}
Z <- rep(0,100)
e <- rnorm(100)
Z[1:5] <- c(3,10,5,0.1,3.5)
for (i in 6:100)   Z[i] <- Z[i-1]+Z[i-4]-Z[i-5]+e[i]-0.4*e[i-1]-0.6*e[i-4]+0.24*e[i-5]

plot(Z,type="l")
Y <- Z[6:100]-Z[5:99]-Z[2:96]+Z[1:95]
plot(Y,type="l")

```

I saw a method to simulate data for a known dist'n([source](https://robjhyndman.com/hyndsight/simulating-from-a-specified-seasonal-arima-model/)), let's try:
```{r}
library(forecast)
set.seed(1)
model <- Arima(ts(rnorm(24000),freq=12), order=c(0,1,1), seasonal=c(0,1,1),fixed=c(theta=0.313, Theta=0.817))
foo <- simulate(model,nsim = 240)
plot(foo,type="l")

foo
summary(model)
```

**Comment: ** our final SARIMA model is $$(1-B)(1-B^{12})Z_t=(1-0.5B)(1-0.5B^{12})\epsilon_t$$. If we take one unit as one year(12 observations), then we have ten years' data.

**---UPDATE 2019.5.26---** I am trying to check the method from the [answer](https://stackoverflow.com/questions/20273104/simulating-a-basic-sarima-model-in-r), but it doesn't work...

```
# install.packages("devtools")
library("devtools")
devtools::install_github("smac-group/gmwm")

# Set seed for reproducibility
set.seed(1)

# Specify a SARIMA(0,1,1)(0,1,1)[12] 
mod = SARIMA(i=1, ma=.5, si = 1, sma = .5, s = 12, sigma2 = 1.5)

# Generate the data
xt2 = gen.gts(mod, 1e3)

# Validate output
arima(xt2, order=c(0,1,1), seasonal=list(order=c(0,1,1), period = 12))
```
## end(perhaps ?)
我感觉数据这块到这这儿就差不多了吧(**too young too naive**)，虽然最后*SARIMA*花了很长时间，走了很多弯路，而且最后的数据我现在还不是很确定能不能用，但是也只能先暂且相信网上的大牛们和自己的判断了。前路茫茫啊，年轻人，不要气馁，继续努力！感觉不能一直给自己说慢慢来，因为感觉目前自己的节奏真的有点太悠闲自在了些...不管怎样，还是要相信自己，坚持你的梦想，朝着梦想前进！别人能做到，为什么我不能呢！多思考，年轻人 :) 
Cheers ~

