---
title: "meeting report 6.10"
author: "linyiguo"
date: "2019.6.10"
output: word_document
---

Well, I did not send Aaron meeting report this time, cause I did not read much things recently, all content we talked was the things I have done before. And the meeting result compared with the former was better. I guess Aaron and I are getting used with each other? LOL.   
I hope to plot something to clearify the point I want to express, but unfortunatly I can't. I will add some photos in this folder.    

# I
Note the curve of *trend + season* in **report 6.6** is pretty close to our data, which is totally opposite to the situation from StatCan... although I am not sure how they get the plot, at least I believe what I got is reasonable.   
Here is the thing, $Y_t-T_t-S_t$ is $\epsilon_t$, and differences between our data and trend+season actually reflect the deviation of the noise $\epsilon$. By controling the deviation of the noise, we can control the smoothness of the trend+season curve(since t.s $Y_t$ is fixed).   


-------------------------------------------**UPDATE 6.12**-------------------------------------------   
昨天试着理清为什么Aaron说可以change $\omega_t=\phi(Y_t-T_t-S_t)$ into $\omega_t=\phi(\frac{Y_t-T_t-S_t}{c}) (c>1)$. But I still do not understand the theory behind it clearly... I give up temporarily :) Let's see whether this could work in our s.s model before(*report 6.6*) 
## change sd in expression of w from 1 to 5 
```{r}
library(seasonal)
library(forecast)
set.seed(1)

# generate data
model <- Arima(ts(rnorm(24000),freq=12), order=c(0,1,1), seasonal=c(0,1,1),fixed=c(theta=0.5, Theta=0.5))
data <- simulate(model,nsim=240)

# define m_x11
m_x11 <- seas(data, x11 = "", regression.aictest =  NULL)

# Initialization
S <- matrix(0,1000,251) # Cause Seasonal component's state space model, we have additional 11 zero-values.
Tr <- matrix(0,1000,251)
Tr[,11] <- data[1]
S[,1:11] <- rep(as.numeric(data-final(m_x11))[1:11],1000)
component <- c()
a = 222 # a-11 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (i in 12:a) {
  
  # update particles
  Tr[,i] <- Tr[,i-1] + rnorm(1000,sd=5)
  for (j in 1:11) S[,i] <- S[,i]-S[,i-j]
  S[,i] <- S[,i] + rnorm(1000, sd=5)
  
  # update weights
  w <- dnorm(data[i-11]-Tr[,i]-S[,i],sd=5)
  w <- w/sum(w)
  
  # evaluate state value
  t <- sum(w * Tr[,i])
  s <- sum(w * S[,i])
  
  # add to our component path
  component <- rbind(component, c(t,s))
  
  # resample
  Tr[,i] <- sample(Tr[,i], size =1000, replace = TRUE, prob = w)
  S[,i] <- sample(S[,i], size = 1000, replace = TRUE, prob = w)
}

# plot four curves together
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,1],type="l",col="green",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,2],type="l",col="blue",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(data[1:(a-11)]-component[,2],type="l",col="red",ylim=c(lower,upper),ylab='')
legend("topleft",c("data","trend","seasonal","deseasoned"),col=c("black","green","blue","red"),lty=c(1,1,1,1))

# data vs deseasoned
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper))
par(new=TRUE)
plot(data[1:(a-11)]-component[,2],type="l",col="red",ylim=c(lower,upper),ylab='')
legend("topleft", c("data","deseasoned"),col=c("black","red"),lty=c(1,1))

# data vs trend
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,1],type="l",col="green",ylim = c(lower,upper),ylab='')
legend("topleft",c("data","trend"),col=c("black","green"),lty=c(1,1))

# data vs season
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,2],type="l",col="blue",ylim = c(lower,upper),ylab='')
legend("topleft",c("data","seasonal"),col=c("black","blue"),lty=c(1,1))

# data vs deseason&detrend
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(data[1:(a-11)]-component[,1]-component[,2], type="l",col="red",ylim = c(lower,upper),ylab='')
legend("topleft",c("data","deseason&detrend"),col=c("black","red"),lty=c(1,1))

# data vs seanson+trend
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,1]+component[,2], type="l",col="red",ylim = c(lower,upper),ylab='')
legend("topleft",c("data","season+trend"),col=c("black","red"),lty=c(1,1))
```


--------------------------------------------**UPDATE 6.16**--------------------------------------------   
I am busy with my resume and the self-recommendation letter recently... I need to meet aaron tmr, but for now I don't have any things to show him, which is very very awkward...   

## sd=100 in weight expression
```{r}
# Initialization
S <- matrix(0,1000,251) # Cause Seasonal component's state space model, we have additional 11 zero-values.
Tr <- matrix(0,1000,251)
Tr[,11] <- data[1]
S[,1:11] <- rep(as.numeric(data-final(m_x11))[1:11],1000)
component <- c()
a = 222 # a-11 is the length of our t.s.
lower = -50  # lower bound
upper = data[a]  # upper bound

for (i in 12:a) {
  
  # update particles
  Tr[,i] <- Tr[,i-1] + rnorm(1000,sd=5)
  for (j in 1:11) S[,i] <- S[,i]-S[,i-j]
  S[,i] <- S[,i] + rnorm(1000,sd=5)
  
  # update weights
  w <- dnorm(data[i-11]-Tr[,i]-S[,i],sd=100)
  w <- w/sum(w)
  
  # evaluate state value
  t <- sum(w * Tr[,i])
  s <- sum(w * S[,i])
  
  # add to our component path
  component <- rbind(component, c(t,s))
  
  # resample
  Tr[,i] <- sample(Tr[,i], size =1000, replace = TRUE, prob = w)
  S[,i] <- sample(S[,i], size = 1000, replace = TRUE, prob = w)
}

# plot four curves together
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,1],type="l",col="green",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,2],type="l",col="blue",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(data[1:(a-11)]-component[,2],type="l",col="red",ylim=c(lower,upper),ylab='')
legend("topleft",c("data","trend","seasonal","deseasoned"),col=c("black","green","blue","red"),lty=c(1,1,1,1))

# data vs deseasoned
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper))
par(new=TRUE)
plot(data[1:(a-11)]-component[,2],type="l",col="red",ylim=c(lower,upper),ylab='')
legend("topleft", c("data","deseasoned"),col=c("black","red"),lty=c(1,1))

# data vs trend
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,1],type="l",col="green",ylim = c(lower,upper),ylab='')
legend("topleft",c("data","trend"),col=c("black","green"),lty=c(1,1))

# data vs season
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,2],type="l",col="blue",ylim = c(lower,upper),ylab='')
legend("topleft",c("data","seasonal"),col=c("black","blue"),lty=c(1,1))

# data vs deseason&detrend
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(data[1:(a-11)]-component[,1]-component[,2], type="l",col="red",ylim = c(lower,upper),ylab='')
legend("topleft",c("data","deseason&detrend"),col=c("black","red"),lty=c(1,1))

# data vs seanson+trend
plot(data[1:(a-11)],type = "l",ylim = c(lower,upper),ylab='')
par(new=TRUE)
plot(component[,1]+component[,2], type="l",col="red",ylim = c(lower,upper),ylab='')
legend("topleft",c("data","season+trend"),col=c("black","red"),lty=c(1,1))
```

Well, after enlarging the deviation of noise, the curves of trend and deseasonal become much more smooth, which should be something like that from StatCan. 

I think I should understand a little bit about why aaron told me to add a demoninator in weights expression. Like we said before, we can control the smoothness of our curves by controling the deviation of our noise.
Let's say we want $$\epsilon \sim N(0,25)$$
which means $$\frac{\epsilon}{5}\sim N(0,1)$$
or $$\frac{Y_t-T_t-S_t}{5} \sim N(0,1)$$
and $$\omega_t \varpropto p(Y_t|T_t,S_t)$$
that is $$\omega_t \varpropto \phi(\frac{Y_t-T_t-S_t}{5} )$$
And in R we can express $\phi(\frac{Y_t-T_t-S_t}{5})$ in two ways:    
-  dnorm((Y_t-T_t-S_t)\5)     
-  dnrom(Y_t-T_t-S_t, sd=5)       
We choose the second one.








